### 2021-10-29 eda/amicus-word-count.Rmd
word count + shortening

---
title: "Amicus Word Count"
author: "Amanda Konet"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F, error = F)

# Box access
library(boxr)
box_auth()

# data manip
library(tidyverse)
library(tidytext)
```

# Purpose

Get word count of amicus briefs -- min, max, med, avg. May need to use long transformers instead of regular since we have some pretty long docs?

consult w/ JSS

```{r}
# raw amicus text
amicus <- box_read("863216279471")

amicus %>% head(5)
```

# Tokenize

Tokenize text as is then get # words per brief

```{r}
# returns case, brief, word for all words in all cases
amicus_tokens <- amicus %>% 
  select(case, brief, text = txt_full) %>%
  group_by(brief) %>% 
  unnest_tokens(word, text)

# word ct per brief
amicus_wrdct <- amicus_tokens %>% group_by(brief) %>% summarize(n=n())
```

# Explore word count

Summary - way higher than BERT.

```{r}
summary(amicus_wrdct$n)
```

LongFormers can use up to 16k tokens... any texts that are longer than this? Yes. We have 45 amicus briefs with 16k+ tokens. Possible we may be able to pare these down but it will have to be manual and we need to make sure we aren't losing information.

```{r}
amicus_wrdct %>% filter(n >= 16000) %>% arrange(-n)
```