### 2021-11-04 data-cleaning/01-shorten-amicus-text.Rmd
reorg

---
title: "Amicus Word Count"
author: "Amanda Konet"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F, error = F)

# Box access
library(boxr)
box_auth()

# data manip
library(tidyverse)
library(tidytext)
```

# Purpose

To pare down amicus text by removing unnecessary/irrelevant text.

1) Remove text AFTER "Summary of Argument" keyword.

2) Remove text AFTER "Appendix" keyword

* "footnotes" ??

Use footnotes when present. If not footnotes, use appendix. If neither: check. There are 154 docs with neither

```{r}
amicus <- amicus %>%
  mutate(txt_full = tolower(txt_full),
         txt_full = str_squish(txt_full)) %>% 
  mutate(appendix = str_count(txt_full, "appendix"),
         footnotes = str_count(txt_full, "footnote"))

amicus %>% select(case, brief, appendix, footnotes) %>% 
  group_by(appendix > 0, footnotes >0) %>% 
  summarize(n())
```

```{r}
amicus %>% filter()
```