### 2022-02-13 eda/calculate_label_density.Rmd
code for #10

---
title: "Create Label Density"
author: "Sarah Torrence"
date: "`r Sys.Date()`"
output: pdf_document
---

The purpose of this script is to find briefs with high label density based on words and phrases that are used frequently in feminist and opposition frames, or arguments. We will do this through a dictionary of frames and the words/phrases associated with those frames in which we will search for these terms within each chunk of text from briefs. Label density will be defined as the total number of times terms associated with each frame is used within the chunk of text.

The end goal is to find the top 5 text chunks with the highest label density for each frame.

## Load Packages

```{r include=FALSE}
# Load packages 
library(boxr) # accessing box API
library(tidyverse) # reading data
library(quanteda) # text mining
box_auth()
```

## Load Data

```{r}
amicus_chunks <- box_read("915435467229")
frames <- box_read("914901620274")
```

## Create Dictionary

We are only focusing on the following frames so I will filter out the rest.

* feminist briefs ("fem" in brief title): 
    + undue burden (law_undueBurdenII)
    + women's rights (law_womanRights)
    + women's citizenship (womanCitizenship)
    + compulsory motherhood (compulsoryMotherhood)
* opposition briefs: 
    + morality (morality)
    + abortion negative (abortionNegativeII)
    + psychological harm (psy_harm)
    + fetus (actor_fetus)
* both types: 
    + evidence (evidence_combine)
    + health (health)

We want to separate these into a fem dictionary and opp dictionary as we only want to search for fem terms in fem briefs and opp terms in opp briefs. We will include `evidence_combine` and `health` in both dictionaries as we want to search for related terms in both fem and opp briefs.

```{r}
#frames to keep
fem_frames <- c('law_undueBurdenII', 'law_womanRights', 'womanCitizenship', 'compulsoryMotherhood', 'evidence_combine', 'health')

opp_frames <- c('morality', 'abortionNegativeII', 'psy_harm', 'actor_fetus', 'evidence_combine', 'health')

#filter for only these frames
fem_df <- frames %>% filter(code %in% fem_frames) %>% select(code, word_cleaned) %>% unique()

opp_df <- frames %>% filter(code %in% opp_frames) %>% select(code, word_cleaned) %>% unique()
```

Here are a couple functions we will use to create dictionaries with frames as keys and associated terms as values.

```{r}
# Get terms associated with each frames
getFrames <- function(frame){ 
  frames %>%
    filter(code == frame) %>%
    select(word_cleaned)
    }

#create dictionary
create_dict <- function(frame_df){

    # Initialize list
    searchFrames <- list()

    # Get unique frames
    frames <- unique(frame_df$code)

    # Pull words for each frame and save in list
    for(i in 1:length(frames)){ 
      searchFrames[i] <- getFrames(frames[i])
    }

    # Name items in list
    names(searchFrames) <- frames

    frameDict <- dictionary(searchFrames)
    return(frameDict)
}
```

Using the above functions to create a fem and opp dictionary.

```{r}
fem_dict <- create_dict(fem_df)
opp_dict <- create_dict(opp_df)
```

## Create Corpus

Now that we have dictionaries of search terms, we need to create a corpus to search through for fem and opp.

```{r}
#function to create a corpus
create_corpus <- function(df){

    brief_corp <- df$text %>%
      corpus()

    # Add meta data to corpus
    docvars(brief_corp, "brief_v1") <- df$V1 
    docvars(brief_corp, "brief_case") <- df$case
    docvars(brief_corp, "brief_id") <- df$id
    docvars(brief_corp, "brief_name") <- df$brief

    # Name documents in corpus
    docnames(brief_corp) <- df$V1
    return(brief_corp)
}

#creating the fem and opp corpus
fem_chunks <- amicus_chunks %>% filter(fem_brief == 1)
fem_corp <- create_corpus(fem_chunks)

opp_chunks <- amicus_chunks %>% filter(fem_brief == 0)
opp_corp <- create_corpus(opp_chunks)
```

Now we can tokenize each corpus. Removing punctuation and numbers will increase the number of term counts (something Amanda previously discovered).

```{r}
# Tokenize brief chunks
fem_toks <- tokens(fem_corp, remove_punct = TRUE, remove_numbers = TRUE)
opp_toks <- tokens(opp_corp, remove_punct = TRUE, remove_numbers = TRUE)
```

Here we can create a summary data frame of each chunk where we will add the frequency of each frame to later.

```{r}
create_summary <- function(corp, chunks){
    # Generate summary of corpus
    corp_summary <- summary(object = corp, 
                          n = nrow(chunks), # default is 100, we need all observations 
                          showmeta = TRUE) %>% # add meta data to summary
      as.data.frame()
    return(corp_summary)
}

#creating summary for fem and opp
fem_summary <- create_summary(fem_corp, fem_chunks)
opp_summary <- create_summary(opp_corp, opp_chunks)
```

## Generate Frame Frequencies

Now we can use the dictionary and tokens to find the frequency of each frame within each chunk of text and add all this information together with the summary information above in a clean data frame.

```{r}
#get the frequency of each frame and add to summary data frame
get_freqs <- function(tokens, dict, summary_df){
        #searching chunks for frames
        df <- tokens_lookup(tokens, dictionary = dict, nested_scope = "dictionary",
                              case_insensitive = FALSE, valuetype = "fixed") %>% dfm()

        #creating data frame of summary info and frequencies
        df <- df %>% convert(to = "data.frame") %>% 
            rename(brief_v1 = doc_id) %>% 
            left_join(summary_df, by = "brief_v1") %>% 
            select(-Text, -Types, -Tokens, -Sentences)
        return(df)
}

#convert v1 to character for joining purposes
fem_summary$brief_v1 <- as.character(fem_summary$brief_v1)
opp_summary$brief_v1 <- as.character(opp_summary$brief_v1)

#call frequency function on fem and opp
fem <- get_freqs(fem_toks, fem_dict, fem_summary)
opp <- get_freqs(opp_toks, opp_dict, opp_summary)
```

The final step is to add back fem and opp into one large data frame with all chunks of text and label density.

```{r}
df_frames_freq <- bind_rows(fem, opp)

amicus_chunks$V1 <- as.character(amicus_chunks$V1)

df_frames_freq <- df_frames_freq %>% left_join(amicus_chunks, by = c('brief_v1' = 'V1')) %>% 
    select(V1 = brief_v1, case, brief_name, id, fem_brief, text, compulsorymotherhood, evidence_combine, health, law_undueburdenii, law_womanrights, womancitizenship, abortionnegativeii, actor_fetus, morality, psy_harm)
```

We also want a boolean value of whether a frame was used at all in each chunk. If a frame was used we will encode a value of 1, if not, a value of 0.

```{r}
create_bool <- function(df, variables){
    for (var in variables){
        df[paste0(var,'_bool')] <- ifelse(df[var] > 0, 1, 0)
    }
    return(df)
}

v <- c('compulsorymotherhood', 'evidence_combine', 'health', 'law_undueburdenii',
       'law_womanrights', 'womancitizenship', 'abortionnegativeii', 'actor_fetus', 
       'morality', 'psy_harm')
df_frames_freq <- create_bool(df_frames_freq, v)
```

## Saving results

```{r}
box_write(df_frames_freq, dir_id = "154895322313", "amicus_chunks_frames_density.csv") 
```

# Finding High Label Density Chunks

Now we can find the top 5 chunks with the highest label density for each of the frames. If

```{r}
#create data frame with 0 rows and 26 columns
df <- data.frame(matrix(ncol = 26, nrow = 0))

#provide column names
colnames(df) <- c("V1","case","brief_name","id","fem_brief","text","compulsorymotherhood",
                  "evidence_combine","health","law_undueburdenii","law_womanrights",
                  "womancitizenship","abortionnegativeii","actor_fetus","morality",
                  "psy_harm","compulsorymotherhood_bool", "evidence_combine_bool",
                  "health_bool","law_undueburdenii_bool", "law_womanrights_bool",
                  "womancitizenship_bool", "abortionnegativeii_bool", "actor_fetus_bool",
                  "morality_bool","psy_harm_bool")

df_na <- df_frames_freq
df_na[is.na(df_na)] <- 0
for (var in v){
    x <- tail(df_na[order(df_na[var]),var],5) %>% unique
    for (i in 1:length(x)){
        y <- df_frames_freq[df_frames_freq[var] == x[i],]
        df <- rbind(df, y)
    }
}

df <- df %>% unique()
```

## Save the Results

```{r}
box_write(df, dir_id = "154895322313", "highest_label_density_chunks.csv")
```