### 2022-02-14 reference-files/search-briefs.Rmd
rename

---
title: "Amicus II: Searching briefs"
output: html_document
---

# Load packages

```{r include=FALSE}
# Load packages 
library(boxr) # accessing box API
library(tidyverse) # reading data
library(quanteda) # text mining
box_auth()
```

# Load Data 
```{r}
briefTextDF <- read_csv("../-processed-data/stop_words_removed/a2-step-2-preprocess-data-brief-text-cleaned.csv")
searchTerms <- read_csv("../-processed-data/stop_words_removed/a2-step-2-preprocess-search-terms-cleaned.csv")
```

# Duplicate Brief

Filter out duplicate June v. Russo Briefs

```{r}
briefTextDF <- briefTextDF %>% 
  filter(!grepl("2,556", ignore.case = T, brief_name))
```

Ensure incorrect use of fact(s), report(s/ing/ed)

```{r}
briefTextDF <- briefTextDF %>%
  mutate(
    brief_text_clean = case_when(
      brief_case_clean == "City of Akron v Akron Center for Rep Health" ~ gsub(
        pattern = " fact | fact\"|\"fact\"|facts",
        replacement = "",
        x = brief_text_clean,
        ignore.case = T
      ),
      brief_case_clean %in% c(
        "Planned Parenthood v Ashcroft",
        "Thornburgh v American Council of OBGYNs"
      ) ~
        gsub(
          pattern = "report|reports|reporting|reported",
          replacement = "",
          x = brief_text_clean,
          ignore.case = T
        ),
      TRUE ~ brief_text_clean
    )
  )
```

# Tokenizing and weighting terms

## Create corpus

In order to generate the frame frequencies, we'll begin by creating a corpus, `amicusCorp`. The corpus is like a file cabinent - it contains all of the texts in our dataset. The one we're creating below contains the preprocessed amicus brief text, `briefTextDF$brief_text_clean`.

Just like when you file away documents in real life, it's important to label items so that you can easily group and categorize them during your analysis. In text mining, these labels/variables are referred to as **meta data**. We'll label the texts in our corpus using the `docvars()` and `docnames()` functions.

Another way to understand how these functions work is imagining that we're moving all of the important variables in our original data frame, `briefTextDF`, to the corpus, `briefCorp`.

```{r}
brief_corp <- briefTextDF$brief_text_clean %>%
  corpus()

# Add meta data to corpus
docvars(brief_corp, "brief_case") <- briefTextDF$brief_case_clean # case name cleaned
docvars(brief_corp, "brief_name") <- briefTextDF$brief_name # brief
docvars(brief_corp, "brief_id") <- briefTextDF$brief_id # file id
docvars(brief_corp, "brief_type") <- briefTextDF$brief_type # party vs amicus
docvars(brief_corp, "brief_party") <- ifelse(briefTextDF$brief_fem == TRUE, "fem", "opp") # feminist or opponent

# Name documents in corpus
docnames(brief_corp) <- briefTextDF$brief_name

summary(brief_corp) %>% head()
```

## Create frame dictionary

This dictionary will be used to filter out frame words from paragraphs when counting word frequency.

```{r}
# Remove duplicates
searchTerms <- searchTerms %>%
  mutate(word = word_cleaned) %>%
  select(abb, word) %>%
  unique()

# Get words associated with frames
getFrames <- function(frame){ 
  searchTerms %>%
    filter(abb == frame) %>%
    select(word)
  }

# Initialize list
searchFrames <- list()

# Get unique frames
frames <- unique(searchTerms$abb)

# Pull words for each frame and save in list
for(i in 1:length(frames)){ 
  searchFrames[i] <- getFrames(frames[i])
}

# Name items in list
names(searchFrames) <- frames

frameDict <- dictionary(searchFrames)

# Check output
frameDict
```

# Look up frames

**Note on arguments in `tokens()` function**: removing punctuation and numbers from terms increases counts. See `amicus-exploring-counting-frames` script to go over evidence.

```{r}
# Tokenize briefs
toks <- tokens(brief_corp, remove_punct = TRUE, remove_numbers = TRUE)
```

## Create brief summary statistics

The `brief_corp_summary` data frame contains the summary statistics of all the briefs in our corpus, `brief_corp`. The one thing it's missing is the number of tokens (i.e. words) in each brief.

The `summary()` function takes 3 arguments:

* **object**: the corpus object (`brief_corp`)
* **n**: the number of observations; the default is 100, but since we want the summary for all of the texts in our corpus, we'll set the value to `nrow(briefTextDF)` since the number of rows (calculated using the function `nrow()`) in the `briefTextDF` is equal to the total number of observations (i.e. briefs) in our corpus.
* **showmeta**: we'll set this value equal to TRUE in order to include the metadata we created earlier.

You'll notice that the `brief_corp` data frame has a column `Tokens` - it's the number of tokens (i.e. words) **before** removing punctuation and digits from the briefs. We'll compare these values to the values we create in the `word_ct` column in a few.

```{r}
# Generate summary of amicusCorp
brief_corp_summary <- summary(object = brief_corp, 
                      n = nrow(briefTextDF), # default is 100, we need all observations 
                      showmeta = TRUE) %>% # add meta data to summary
  as.data.frame()

# Look at first 6 observations 
tail(brief_corp_summary)
```

Now it's time to find out how many words are in the briefs after we tokenize them and remove the punctuation and digits. We can do this by using the `ntokens()` function - this function gets the counts of tokens (total words/features) in our `toks` object.

The `nToks` data frame contains to columns: 
* **brief_name**: the name of the brief associated with the `ntoken(toks)` object (i.e. value in `brief_n_words`); this column allows us to join the data frame with the `amicusCorpSummary` data frame. 
* **brief_n_words**: the number of tokens (i.e. words) in each brief **after** removing the punctuation and digits from the briefs.

We'll join this information to `amicusCorpSummary` to create the data frame `briefInfo`.

```{r}
# Total number of tokens (before pulling frame words)
nToks <- data.frame(brief_name = names(ntoken(toks)), 
                     word_ct = ntoken(toks))

# Look at first 6 observations
head(nToks)
```

```{r}
brief_info <- brief_corp_summary %>%
  left_join(nToks, by = c("brief_name")) %>%
  select(brief_case, brief_name, brief_party, brief_type, word_ct, brief_id)

head(brief_info)
```

### OPTIONAL: Comparing word count before/after tokenizing briefs

We can compare the word count before removing the puncutation and digits to the word count after removing the puncutation and digits by running the code chunk below.

```{r}
# Compare word counts
# compare_brief_n_words <- amicusCorpSummary %>%
#  select(brief_name, 
#         brief_n_words_raw = Tokens) %>% # number of tokens before removing punctuation and digits 
#  left_join(nToks, by = c("brief_name")) %>% 
#  select(brief_name, brief_n_words_raw, brief_n_words) %>%
#  mutate(same_number = brief_n_words == brief_n_words_raw) # are word counts the same
```

If the value in the column `same_number` is TRUE, then the brief word count did not change after removing the punctuation and digits. 
If the value in the column `same_number` is FALSE, then the brief word count changed (it decreased) after removing the punctuation and digits.

We can use the `count(same_number)` to count the number of observations that are TRUE and FALSE in the `same_number` column, and based on the output, all 670 word counts changed after removing the punctuation and digits.

```{r}
# Did any word counts stay the same? Nope
# compare_brief_n_words %>%
#   count(same_number)
```

Taking it one more step, we can look at how the word count values changed - we'd expect them to decrease after removing the puncutation and digits, but let's double check by running one more test:

```{r}
# Did any word counts increase after removing punctuation and digits? Nope (whew!)
# compare_brief_n_words %>%
#   mutate(greater_number = brief_n_words > brief_n_words_raw) %>% 
#   filter(greater_number == TRUE)
```

## Generate frame frequencies

Now that we've created our `brief_info` data frame, it's time to generate the frame frequencies and attach the data frames together.

The `brief_frames_raw` data frame contains the raw frame frequencies joined with the `brief_info` data frame. The resulting dataframe of this step will be one row per case per frame, with a column for the count of that frame. This will make it very easy to generate the frame frequencies per Holly's specifications, which is:

(# of words/phrases for a particular frame) / (# of all words in the brief)

```{r}
# Search for frames
brief_dfm <- tokens_lookup(toks, dictionary = frameDict, nested_scope = "dictionary", case_insensitive = FALSE, valuetype = "fixed") %>%
  dfm()

# Generate raw frequencies
brief_frames_raw <- brief_dfm %>%
  convert(to = "data.frame") %>%
  rename(brief_name = doc_id) %>%
  left_join(brief_info, by = "brief_name") %>% # join with brief_info
  # remove "purpose of performance" counts from "purpose" counts
  mutate(p_new = p - pop) %>% 
  select(brief_case, brief_id, brief_name, brief_type, brief_party, word_ct, an:wc)

# Check out first 6 observations 
head(brief_frames_raw, 10)
```
**Update 4/22/21**: Remove instances of word usage from the evidence experts frame that we don't want to count.

For Thornburgh, remove counts of the words "fact(s)" and "evidence" from the opponent/antiabortion party brief. Per github issue #92, 6+14 instances of the eeII frame must be subtracted

```{r}
brief_frames_raw <- brief_frames_raw %>%
  mutate(eeii = ifelse(
    grepl("Thornburgh", brief_case) &
      brief_type == "Party" & brief_party == "opp",
    eeii - (6 + 14),
    eeii
  ))

#brief_frames_raw %>%
#  filter(grepl("Thornburgh", brief_case), brief_type == "Party", brief_party == "opp") %>% 
#  select(brief_case, brief_type, brief_party, eeii)
```

For City of Akron, remove the counts of "fact" from select amicus (opp + fem briefs) found in step 3b.

```{r}
akron_cases <- c("City of Akron v Akron Center for Rep Health. Amici Brief for Petitioner, by United Families Foundation and Women Exploited.docx", "City of Akron v Akron Center for Rep Health. Amici Brief for Respondent (feminist), by American College of OBGYNs et al.docx", "City of Akron v Akron Center for Rep Health. Amici Brief for Respondent (feminist), by Planned Parenthood Federation of America et al.docx", "City of Akron v Akron Center for Rep Health. Amicus Brief for Petitioner, by Catholic League for Religious and Civil Rights.docx", "City of Akron v Akron Center for Rep Health. Amicus Brief for Petitioner, by Feminists for Life.docx", "City of Akron v Akron Center for Rep Health. Amicus Brief for Petitioner, by Legal Defense Fund for Unborn Children.docx", "City of Akron v Akron Center for Rep Health. Amicus Brief for Petitioner, by Womenkind, Inc.docx", "City of Akron v Akron Center for Rep Health. Amicus Brief for Respondent (feminist), by Certain Religious Organizations.docx", "City of Akron v Akron Center for Rep Health. Amicus Brief for Respondent (feminist), by the American Public Health Association.docx")

#brief_frames_raw %>% 
#  filter(brief_name %in% akron_cases) %>% 
#  select(brief_name, eeii)

brief_frames_raw <- brief_frames_raw %>% 
  mutate(eeii = case_when(
    brief_name == "City of Akron v Akron Center for Rep Health. Amici Brief for Petitioner, by United Families Foundation and Women Exploited.docx" ~ eeii - 1,
    brief_name == "City of Akron v Akron Center for Rep Health. Amici Brief for Respondent (feminist), by American College of OBGYNs et al.docx" ~ eeii - 2,
    brief_name == "City of Akron v Akron Center for Rep Health. Amici Brief for Respondent (feminist), by Planned Parenthood Federation of America et al.docx" ~ eeii - 2,
    brief_name == "City of Akron v Akron Center for Rep Health. Amicus Brief for Petitioner, by Catholic League for Religious and Civil Rights.docx" ~ eeii - 2,
    brief_name == "City of Akron v Akron Center for Rep Health. Amicus Brief for Petitioner, by Feminists for Life.docx" ~ eeii - 2,
    brief_name == "City of Akron v Akron Center for Rep Health. Amicus Brief for Petitioner, by Legal Defense Fund for Unborn Children.docx" ~ eeii - 6,
    brief_name == "City of Akron v Akron Center for Rep Health. Amicus Brief for Petitioner, by Womenkind, Inc.docx" ~ eeii - 1,
    brief_name == "City of Akron v Akron Center for Rep Health. Amicus Brief for Respondent (feminist), by Certain Religious Organizations.docx" ~ eeii - 1,
    brief_name == "City of Akron v Akron Center for Rep Health. Amicus Brief for Respondent (feminist), by the American Public Health Association.docx" ~ eeii - 1,
    T ~ eeii
  ))
```

Get frame frequencies

```{r}
# Generate frame frequencies 
brief_frames_freq <- brief_frames_raw %>% 
  pivot_longer(c(an:wc), names_to = "frame", values_to = "frame_ct") %>% 
  mutate(frame_freq = frame_ct/word_ct) %>% 
  select(-frame_ct) %>% 
  pivot_wider(names_from = "frame", values_from = frame_freq)

brief_frames_freq %>% head(10)
```

## Raw counts + freq

Dataframe with raw counts and frequencies

```{r}
brief_frames <- brief_frames_raw %>% 
  left_join(., brief_frames_freq, 
            by = c("brief_case", "brief_id", "brief_name", "brief_type", "brief_party", "word_ct"), 
            suffix = c("_raw", "_freq")) %>% 
  select(brief_case:word_ct, sort(colnames(.)))

brief_frames %>% head(10)
```

Summary by case

```{r}
# Total word count of briefs by case
brief_frames %>%
  group_by(brief_case) %>%
  summarise(total_brief_words = sum(word_ct)) 
```

## Save results

```{r}
# Save files
write_csv(brief_frames_raw, "../-processed-data/stop_words_removed/a2-step-3-search-briefs-raw-freq.csv")
write_csv(brief_frames_freq,"../-processed-data/stop_words_removed/a2-step-3-search-briefs-ter-freq.csv")
write_csv(brief_frames, "../-processed-data/stop_words_removed/a2-step-3-search-briefs-full-freq.csv")

# all stops removed folder
#box_write(brief_frames, dir_id = "126163615802", "search-briefs-full-freq.csv")

# group a stops folder
box_write(brief_frames, dir_id = "126482779345", "search-briefs-full-freq.csv")

rm(list = ls())
```