### 2022-04-02 data-cleaning/04-create-walkover-brief-name.Rmd
started walkover of brief names

---
title: "Create Walkover of Brief Names"
author: "Sarah Torrence"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F, error = F)

# Box access
library(boxr)
box_auth()

# data manip
library(tidyverse)
library(fuzzyjoin)

# data
# regressions/data for Amicus IV regressions/Amicus Author Organization Categorization
brief_info <- box_read("936715223606")

# data/processed amicus files/amicus_text_512.csv
amicus <- box_read("937207017356")
```

```{r}
brief_names <- brief_info %>% filter(`Row Type...9` == "Amicus Brief") %>% 
              select(`Amicus Brief`) %>% unique()

briefs_to_match <- amicus %>% select(brief) %>% unique()
```

Currently none of the names of briefs match between these two files.

```{r}
intersect(brief_names[["Amicus Brief"]], briefs_to_match[["brief"]])
```

It looks like the brief names in the file with text for topic modeling are a combination of case names and brief names. I want to remove the case portion and see if they match better with the brief names in the author information data.

1. Make everything lower case
2. Replace all v. with v
3. Remove .docx
4. Remove all text at beginning of string with `text v text. ` or `text v text - `
5. Remove `%2c`
6. Remove all punctuation and spacing

```{r}
#1
briefs_to_match$clean_names <- tolower(briefs_to_match$brief)
brief_names$clean_names <- tolower(brief_names$`Amicus Brief`)

#2
briefs_to_match$clean_names <- str_replace(briefs_to_match$clean_names, 'v\\.', 'v')

#3
briefs_to_match$clean_names <- str_replace(briefs_to_match$clean_names, '\\.docx', '')

#4
briefs_to_match$clean_names <- str_replace(briefs_to_match$clean_names, '([^v]*v[^\\.]*\\. )|([^v]*v[^-]*- )', '')

#5
briefs_to_match$clean_names <- str_replace(briefs_to_match$clean_names, '%2c', '')

#6
briefs_to_match$clean_names <- str_replace_all(briefs_to_match$clean_names,
                                               '[^[:alnum:]]','')
brief_names$clean_names <- str_replace_all(brief_names$clean_names,
                                               '[^[:alnum:]]','')
```

Sorting by brief name.

```{r}
brief_names <- brief_names[order(brief_names$clean_names),]
briefs_to_match <- briefs_to_match[order(briefs_to_match$clean_names),]
```

After all of the cleaning I was only able to match 268 brief names.

```{r}
join <- brief_names %>% left_join(briefs_to_match, by = c("clean_names"))

table(is.na(join$brief))
```