### 2022-04-08 data-cleaning/06-add-amicus-measures.Rmd
join segment vars close #15

---
title: "Add Amicus Measures"
author: "Sarah Torrence"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F, error = F)

# Box access
library(boxr)
box_auth()

# data manip
library(tidyverse)
library(janitor)

# data
# regressions/data for Amicus IV regressions/Amicus Author Organization Categorization.xlsx
brief_info <- box_read("936715223606")

# data/processed amicus files/amicus_text_clean_512.csv
amicus <- box_read("939714323391")

# data/processed amicus files/walkover_brief_names.csv
walkover <- box_read("941154488137")
```

We need to grab only the amicus brief information from the `brief_info`file, select the variables we care about and impute NAs with 0.

Variable to select:

forgau, antauo, chcauf, lifauo, govauf, govauo, sg, racauf, racauo, medauf, medauo, mpeauf, mpeauo, legauf, legauo, relauf, relauo, LGBauf, forgauf, aclauf, aclauo, sgauf, sgauo

```{r}
col_name <- names(brief_info)[1:12]
brief_info <- brief_info %>%
  row_to_names(row_number = 1)
colnames(brief_info)[1:12] <- col_name
colnames(brief_info)[177] <- 'emppf2'
colnames(brief_info)[5] <- 'amicus_id'

brief_info <- brief_info %>% filter(`Row Type` == 'Amicus Brief') %>% select(Case, 
              amicus_id, `Amicus Brief`, forgau, antauo, chcauf, lifauo, govauf,                     govauo, sg, racauf, racauo, medauf, medauo, mpeauf, mpeauo, legauf, 
               legauo, relauf, relauo, LGBauf, forgauf, aclauf, aclauo, sgauf, sgauo)

# change june v russo to case id = 999999
brief_info <- brief_info %>%
  mutate(Case = ifelse(is.na(Case), "999999", Case))

#impute all na with 0
brief_info[is.na(brief_info)] = 0
```

Join to amicus data

```{r}
walkover$case_code <-  as.character(walkover$case_code)
walk_measures <- walkover %>% left_join(brief_info, by = c("case_code" = "Case", "brief_code" = "amicus_id"))

amicus <- amicus %>% left_join(walk_measures, by = c("id" = "new_id"))
```

Remove unnecessary columns

```{r}
final <- amicus %>% 
  select(case, id, brief, brief_party, forgau:sgauo, text) %>% 
  janitor::clean_names()
```

Check

```{r}
final %>% head()
```

Save as a csv.

```{r}
box_write(final,
           file_name = "amicus_text_clean_512_vars.csv",
           dir_id = 145990409196)
```