### 2022-04-18 eda/capstone/capstone-report.Rmd
capstone report eda

---
title: "Capstone Graphs"
author: "Amanda Konet"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F, error = F)

# Box access
library(boxr)
box_auth()

# data manip
library(tidyverse)
library(janitor)

# plotting
library(ggplot2)
library(ggpubr)

# data
amicus <- box_read("942861514695")

# brief info + clean 
brief_info <- box_read("936715223606")

col_name <- names(brief_info)[1:12]
brief_info <- brief_info %>%
  row_to_names(row_number = 1)
colnames(brief_info)[1:12] <- col_name
colnames(brief_info)[177] <- 'emppf2'
colnames(brief_info)[5] <- 'amicus_id'

brief_info <- brief_info %>% filter(`Row Type` == 'Amicus Brief') %>% select(Case, 
              amicus_id, `Amicus Brief`, forgau, antauo, chcauf, lifauo, govauf,
              govauo, sg, racauf, racauo, medauf, medauo, mpeauf, mpeauo, legauf, 
              legauo, relauf, relauo, LGBauf, forgauf, aclauf, aclauo, sgauf, sgauo)

# change june v russo to case id = 999999
brief_info <- brief_info %>%
  mutate(Case = ifelse(is.na(Case), "999999", Case))

#impute all na with 0
brief_info[is.na(brief_info)] = 0

medauo <- readr::read_csv("~/Documents/Vanderbilt/SP_22/capstone/medauo_topics_clean_labels_bbu_rramicus.csv",
                          col_types = cols(X1 = col_skip()))

medauf <- readr::read_csv("~/Documents/Vanderbilt/SP_22/capstone/medauf_topics_clean_labels_bbu_rramicus.csv",
           col_types = cols(X1 = col_skip()))
```

# Capstone Graphics

# Amicus author type counts

```{r}
amicus %>% distinct(id, brief_party) %>% group_by(brief_party) %>% summarize(n=n())
```

```{r}
brief_info %>% select(forgau:sgauo) %>%
  mutate(across(where(is.character), as.numeric)) %>% 
  summarise(across(where(is.numeric), ~ sum(.x, na.rm = TRUE))) %>% 
  pivot_longer(., cols = c(forgau:sgauo), names_to = "var", values_to = "ct") %>% 
  arrange(var)
```

# MLM fine-tuning

Side-by-side, one for BERT & one for legalBERT

BERT

```{r}
bert <- data.frame(
  epoch = c(1:10),
  train_loss = c(2.0341, 1.7881, 1.6970, 1.6328, 1.5919, 1.5602, 1.5221, 1.5135, 1.4917, 1.4859),
  val_loss = c(1.7515, 1.6616, 1.6061, 1.5662, 1.5362, 1.5193, 1.4984, 1.4988, 1.4755, 1.4671)
)
bert <- bert %>% mutate(train_perp = exp(train_loss), val_perp = exp(val_loss))
```

legalBERT

```{r}
legalbert <- data.frame(
  epoch = c(1:10),
  train_loss = c(1.021, 1.2272, 1.2467, 1.2149, 1.1855, 1.1674, 1.1508, 1.1386, 1.129, 1.1193),
  val_loss = c(1.3393, 1.2612, 1.2403, 1.2276, 1.2101, 1.2020, 1.893, 1.1870, 1.1794, 1.1759)
)
legalbert <- legalbert %>% 
  mutate(train_perp = exp(train_loss), val_perp = exp(val_loss))
#legalbert
```
Plot

```{r}
colors_line <- c("Train" = "#DC277F", "Validation" = "#658FFF")

bert_train <- bert %>% 
  ggplot(.,aes(x = epoch)) +
  geom_line(aes(y = train_perp, color = "Train")) +
  geom_point(aes(y = train_perp), color = "#DC277F", alpha = 0.8) + 
  geom_line(aes(y = val_perp, color = "Validation")) +
  geom_point(aes(y = val_perp), color = "#658FFF", alpha = 0.8) +
  scale_color_manual(values = colors_line) +
  coord_cartesian(ylim = c(0,8), xlim=c(1,10)) +
  scale_y_continuous(breaks = seq(0, 8, by = 1)) +
  scale_x_continuous(breaks = seq(1, 10, by = 1)) +
  labs(
    x = "Epoch",
    y = "Perplexity",
    color = "Dataset",
    title = "BERT MLM Fine-Tuning"
  ) +
  theme_minimal()

bert_train
```

```{r}
lb_train <- legalbert %>%
  ggplot(.,aes(x = epoch)) +
  geom_line(aes(y = train_perp, color = "Train")) +
  geom_point(aes(y = train_perp), color = "#DC277F", alpha = 0.8) + 
  geom_line(aes(y = val_perp, color = "Validation")) +
  geom_point(aes(y = val_perp), color = "#658FFF", alpha = 0.8) +
  scale_color_manual(values = colors_line) +
  coord_cartesian(ylim = c(0,8), xlim=c(1,10)) +
  scale_y_continuous(breaks = seq(0, 8, by = 1)) +
  scale_x_continuous(breaks = seq(1, 10, by = 1)) +
  labs(
    x = "Epoch",
    y = "Loss",
    color = "Loss Type",
    title = "Legal-BERT MLM Fine-Tuning"
  ) +
  theme_minimal() #+
  #theme(panel.grid.minor.x = element_blank())

lb_train
```

Together

```{r}
trains <- ggarrange(bert_train, lb_train, ncol=2, nrow=1, common.legend = TRUE, legend="bottom")
ggsave(filename = "training_results.png", trains, width = 8, height=3)
```

# Topic Modeling Results

## Medauo

new label
1 = fetus viability
2 = abortion ending human life 
3 = public funds not for abortion
4 = fetus having right to life, autonomy, personhood
5 = women's health / mortality from abortion

```{r}
medauo %>% 
```
Orig

```{r, fig.width=8px, fig.height=1px}
medauo_orig <- medauo %>% 
  ggplot(., aes(x=x, y=y, color=as.factor(label), size=Count)) +
  geom_point(alpha=0.6) +
  theme_minimal() +
  guides(size = "none") +
  labs(
    x = "Dimension 2", y = "Dimension 1",
    title = "Original Medical Org Arguments (opp)",
    color = "Label"
  )
ggsave("medauo_orig.png",medauo_orig,height = 4, width = 4)
getwd()
```

Adjusted

```{r}
medauo <- medauo %>% 
  mutate(label_correct = case_when(
    label == 1 ~ "Abortion Ends Human Life",#2,
    label == 3 & (topic_id %in% c(11,28)) ~ "Public Funds",#3,
    label == 3 & !(topic_id %in% c(11,28)) ~ "Fetus Personhood",#4,
    label == 4 & (topic_id %in% c(2,6)) ~ "Fetus Viability",#1,
    label == 5 ~ "Abortion Harms Health",#5,
    T ~ "none"
  ))
```

```{r}
medauo %>% filter(label_correct == "Abortion Ends Human Life")
```

```{r}
medauo_adj <- medauo %>% 
  filter(label_correct != "none") %>% 
  ggplot(., aes(x=x, y=y, color=as.factor(label_correct), size=Count)) +
  geom_point(alpha=0.6) +
  theme_minimal() +
  guides(size = "none") +
  labs(
    x = "Dimension 2", y = "Dimension 1",
    title = "Adjusted Medical Org Arguments (opp)",
    color = "Argument"
  )
ggsave("medauo_adj.png",medauo_adj,height = 4, width = 6)
medauo_adj
```

```{r}
medauo_graphs <- ggarrange(medauo_orig, medauo_adj, ncol=2, nrow=1, common.legend = FALSE)
ggsave(filename = "medauo_results.png", medauo_graphs, width = 6, height=3)
medauo_graphs
```

## Medauf

```{r}
medauf %>% 
  #filter(label %in% c(4, 8,9)) %>% 
  ggplot(., aes(x=x, y=y, color=as.factor(label), size=Count)) +
  geom_point(alpha=0.6) +
  theme_minimal() +
  guides(size = "none")
```

new labels

removing: 2, 7

1. public funding (medicaid); 13,17, 35, 38, 22; orig -1, 0
2. violence against abortion seekers; orig 1, 11
3. minor's rights; 9, 51, 58, 59; orig 4, 8 (notifying parents as bad, counseling, risks to teenage pregnancies)
4. Physician's admitting privileges; orig 5
5. health of abortion-seekers; orig 6, 12
6. abortion & contraceptives access (orig 9)

```{r}
medauf <- medauf %>% 
  mutate(label_correct = case_when(
    label %in% c(-1, 0) ~ "Support Public Funding (Medicaid)",
    label %in% c(1,11) ~ "Violence Against Abortion Seekers",
    label %in% c(4,8) & !(topic_id %in% c(1,25)) ~ "Minor's Rights",
    label == 5 ~ "Physicians Admitting Privileges",
    label %in% c(6, 12) ~ "Health of Abortion-Seekers",
    label == 9 ~ "Abortion/Contraceptives Access",
    T ~ "none"
  ))
```

New graph

```{r}
medauf_adj <- medauf %>% 
  filter(label_correct != "none") %>% 
  ggplot(., aes(x=x, y=y, color=as.factor(label_correct), size=Count)) +
  geom_point(alpha=0.6) +
  theme_minimal() +
  guides(size = "none") +
  labs(
    x = "Dimension 2", y = "Dimension 1",
    title = "Adjusted Medical Org Arguments (pro)",
    color = "Argument"
  )
medauf_adj
```

```{r}
ggsave("medauf_adj.png",medauf_adj,height = 4, width = 6)
```

```{r}
medauf %>% filter(label_correct == "Health of Abortion-Seekers")
```