### 2022-05-02 eda/abortion-cases-frames-plots.Rmd
misc

---
title: "abortion cases"
author: "Amanda Konet"
date: "`r date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F, error = F)
library(tidyverse)
library(ggplot2)
library(boxr)
box_auth()
```

```{r}
df <- box_read("949904531983", read_fun = readr::read_csv)
```

# Abortion Negative II

```{r}
df %>% 
  select(case, ddate2, ) %>% 
  mutate(title = "Word/Phrase Count for Abortion Negative II Frame",
         year = lubridate::year(decision_date)) %>% 
  ggplot(., aes(x = reorder(case, decision_date), y = anii_freq)) +
  geom_bar(stat = "identity") +
  labs(
    x = "Case",
    y = "Frequency of frame words/phrases in decision (divided by total word count)",
    title = "Word/Phrase Count for Abortion Negative II Frame"
  ) + 
  facet_grid(~ year, switch = "x", scales = "free_x", space = "free_x") +
  theme_bw() +
  theme(
    # grid lines + background
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "white", size = 0.1),
    panel.border = element_blank(),
    plot.background = element_rect(color = "#eaeaea", size = 0.4),
    axis.line = element_line(color = "#eaeaea", size = 0.4),
    # facet format
    panel.spacing = unit(0, "pt"), 
    strip.background = element_blank(),
    strip.placement = "outside",
    strip.text = element_text(color = "#afafaf", size = 7.4),
    # axis title
    axis.title = element_text(color = "#666666", size = 8),
    # axis text
    axis.text.x = element_text(color = "#afafaf", size = 7.4, vjust = 0.9),
    axis.text.y = element_text(color = "darkgray", size = 7),
    # axis ticks
    axis.ticks.x = element_line(colour = "#bdbdbd", size = 0.4),
    axis.ticks.y = element_blank(),
    # title
    plot.title = element_text(color = "#666666", size = 9.1, hjust = 0.5)
  )
```

\newpage

# actor_couple

```{r}
dec_frames %>% 
  select(case, decision_date, ac_freq) %>% 
  mutate(title = "Word/Phrase Count for Actor Couple Frame",
         year = lubridate::year(decision_date)) %>% 
  ggplot(., aes(x = reorder(case, year), y = ac_freq)) +
  geom_bar(stat = "identity") +
  labs(
    x = "Case",
    y = "Frequency of frame words/phrases in decision (divided by total word count)",
    title = "Word/Phrase Count for Actor Couple Frame"
  ) + 
  facet_grid(~ year, switch = "x", scales = "free_x", space = "free_x") +
  theme_bw() +
  theme(
    # grid lines + background
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "white", size = 0.1),
    panel.border = element_blank(),
    plot.background = element_rect(color = "#eaeaea", size = 0.4),
    axis.line = element_line(color = "#eaeaea", size = 0.4),
    # facet format
    panel.spacing = unit(0, "pt"), 
    strip.background = element_blank(),
    strip.placement = "outside",
    strip.text = element_text(color = "#afafaf", size = 7.4),
    # axis title
    axis.title = element_text(color = "#666666", size = 8),
    # axis text
    axis.text.x = element_text(color = "#afafaf", size = 7.4, vjust = 0.9),
    axis.text.y = element_text(color = "darkgray", size = 7),
    # axis ticks
    axis.ticks.x = element_line(colour = "#bdbdbd", size = 0.4),
    axis.ticks.y = element_blank(),
    # title
    plot.title = element_text(color = "#666666", size = 9.1, hjust = 0.5)
  )
```

\newpage

# actor_fetus

```{r}
dec_frames %>% 
  select(case, decision_date, af_freq) %>% 
  mutate(year = lubridate::year(decision_date)) %>% 
  ggplot(., aes(x = reorder(case, year), y = af_freq)) +
  geom_bar(stat = "identity") +
  labs(
    x = "Case",
    y = "Frequency of frame words/phrases in decision (divided by total word count)",
    title = "Word/Phrase Count for Actor Fetus Frame"
  ) + 
  facet_grid(~ year, switch = "x", scales = "free_x", space = "free_x") +
  theme_bw() +
  theme(
    # grid lines + background
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "white", size = 0.1),
    panel.border = element_blank(),
    plot.background = element_rect(color = "#eaeaea", size = 0.4),
    axis.line = element_line(color = "#eaeaea", size = 0.4),
    # facet format
    panel.spacing = unit(0, "pt"), 
    strip.background = element_blank(),
    strip.placement = "outside",
    strip.text = element_text(color = "#afafaf", size = 7.4),
    # axis title
    axis.title = element_text(color = "#666666", size = 8),
    # axis text
    axis.text.x = element_text(color = "#afafaf", size = 7.4, vjust = 0.9),
    axis.text.y = element_text(color = "darkgray", size = 7),
    # axis ticks
    axis.ticks.x = element_line(colour = "#bdbdbd", size = 0.4),
    axis.ticks.y = element_blank(),
    # title
    plot.title = element_text(color = "#666666", size = 9.1, hjust = 0.5)
  )
```

\newpage

# actor_physician

```{r}
dec_frames %>% 
  select(case, decision_date, ap_freq) %>% 
  mutate(year = lubridate::year(decision_date)) %>% 
  ggplot(., aes(x = reorder(case, year), y = ap_freq)) +
  geom_bar(stat = "identity") +
  labs(
    x = "Case",
    y = "Frequency of frame words/phrases in decision (divided by total word count)",
    title = "Word/Phrase Count for Actor Physician Frame"
  ) + 
  facet_grid(~ year, switch = "x", scales = "free_x", space = "free_x") +
  theme_bw() +
  theme(
    # grid lines + background
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "white", size = 0.1),
    panel.border = element_blank(),
    plot.background = element_rect(color = "#eaeaea", size = 0.4),
    axis.line = element_line(color = "#eaeaea", size = 0.4),
    # facet format
    panel.spacing = unit(0, "pt"), 
    strip.background = element_blank(),
    strip.placement = "outside",
    strip.text = element_text(color = "#afafaf", size = 7.4),
    # axis title
    axis.title = element_text(color = "#666666", size = 8),
    # axis text
    axis.text.x = element_text(color = "#afafaf", size = 7.4, vjust = 0.9),
    axis.text.y = element_text(color = "darkgray", size = 7),
    # axis ticks
    axis.ticks.x = element_line(colour = "#bdbdbd", size = 0.4),
    axis.ticks.y = element_blank(),
    # title
    plot.title = element_text(color = "#666666", size = 9.1, hjust = 0.5)
  )
```

\newpage

# compulsoryII

```{r}
dec_frames %>% 
  select(case, decision_date, c2_freq) %>% 
  mutate(year = lubridate::year(decision_date)) %>% 
  ggplot(., aes(x = reorder(case, year), y = c2_freq)) +
  geom_bar(stat = "identity") +
  labs(
    x = "Case",
    y = "Frequency of frame words/phrases in decision (divided by total word count)",
    title = "Word/Phrase Count for Compulsory II Frame"
  ) + 
  facet_grid(~ year, switch = "x", scales = "free_x", space = "free_x") +
  theme_bw() +
  theme(
    # grid lines + background
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "white", size = 0.1),
    panel.border = element_blank(),
    plot.background = element_rect(color = "#eaeaea", size = 0.4),
    axis.line = element_line(color = "#eaeaea", size = 0.4),
    # facet format
    panel.spacing = unit(0, "pt"), 
    strip.background = element_blank(),
    strip.placement = "outside",
    strip.text = element_text(color = "#afafaf", size = 7.4),
    # axis title
    axis.title = element_text(color = "#666666", size = 8),
    # axis text
    axis.text.x = element_text(color = "#afafaf", size = 7.4, vjust = 0.9),
    axis.text.y = element_text(color = "darkgray", size = 7),
    # axis ticks
    axis.ticks.x = element_line(colour = "#bdbdbd", size = 0.4),
    axis.ticks.y = element_blank(),
    # title
    plot.title = element_text(color = "#666666", size = 9.1, hjust = 0.5)
  )
```

\newpage

# compulsoryMotherhood

```{r}
dec_frames %>% 
  select(case, decision_date, cm_freq) %>% 
  mutate(year = lubridate::year(decision_date)) %>% 
  ggplot(., aes(x = reorder(case, year), y = cm_freq)) +
  geom_bar(stat = "identity") +
  labs(
    x = "Case",
    y = "Frequency of frame words/phrases in decision (divided by total word count)",
    title = "Word/Phrase Count for Compulsory Motherhood Frame"
  ) + 
  facet_grid(~ year, switch = "x", scales = "free_x", space = "free_x") +
  theme_bw() +
  theme(
    # grid lines + background
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "white", size = 0.1),
    panel.border = element_blank(),
    plot.background = element_rect(color = "#eaeaea", size = 0.4),
    axis.line = element_line(color = "#eaeaea", size = 0.4),
    # facet format
    panel.spacing = unit(0, "pt"), 
    strip.background = element_blank(),
    strip.placement = "outside",
    strip.text = element_text(color = "#afafaf", size = 7.4),
    # axis title
    axis.title = element_text(color = "#666666", size = 8),
    # axis text
    axis.text.x = element_text(color = "#afafaf", size = 7.4, vjust = 0.9),
    axis.text.y = element_text(color = "darkgray", size = 7),
    # axis ticks
    axis.ticks.x = element_line(colour = "#bdbdbd", size = 0.4),
    axis.ticks.y = element_blank(),
    # title
    plot.title = element_text(color = "#666666", size = 9.1, hjust = 0.5)
  )
```

\newpage

# favorChildbirth

```{r}
dec_frames %>% 
  select(case, decision_date, fc_freq) %>% 
  mutate(year = lubridate::year(decision_date)) %>% 
  ggplot(., aes(x = reorder(case, year), y = fc_freq)) +
  geom_bar(stat = "identity") +
  labs(
    x = "Case",
    y = "Frequency of frame words/phrases in decision (divided by total word count)",
    title = "Word/Phrase Count for Favor Childbirth Frame"
  ) + 
  facet_grid(~ year, switch = "x", scales = "free_x", space = "free_x") +
  theme_bw() +
  theme(
    # grid lines + background
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "white", size = 0.1),
    panel.border = element_blank(),
    plot.background = element_rect(color = "#eaeaea", size = 0.4),
    axis.line = element_line(color = "#eaeaea", size = 0.4),
    # facet format
    panel.spacing = unit(0, "pt"), 
    strip.background = element_blank(),
    strip.placement = "outside",
    strip.text = element_text(color = "#afafaf", size = 7.4),
    # axis title
    axis.title = element_text(color = "#666666", size = 8),
    # axis text
    axis.text.x = element_text(color = "#afafaf", size = 7.4, vjust = 0.9),
    axis.text.y = element_text(color = "darkgray", size = 7),
    # axis ticks
    axis.ticks.x = element_line(colour = "#bdbdbd", size = 0.4),
    axis.ticks.y = element_blank(),
    # title
    plot.title = element_text(color = "#666666", size = 9.1, hjust = 0.5)
  )
```

\newpage

# health

```{r}
dec_frames %>% 
  select(case, decision_date, he_freq) %>% 
  mutate(year = lubridate::year(decision_date)) %>% 
  ggplot(., aes(x = reorder(case, year), y = he_freq)) +
  geom_bar(stat = "identity") +
  labs(
    x = "Case",
    y = "Frequency of frame words/phrases in decision (divided by total word count)",
    title = "Word/Phrase Count for Health Frame"
  ) + 
  facet_grid(~ year, switch = "x", scales = "free_x", space = "free_x") +
  theme_bw() +
  theme(
    # grid lines + background
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "white", size = 0.1),
    panel.border = element_blank(),
    plot.background = element_rect(color = "#eaeaea", size = 0.4),
    axis.line = element_line(color = "#eaeaea", size = 0.4),
    # facet format
    panel.spacing = unit(0, "pt"), 
    strip.background = element_blank(),
    strip.placement = "outside",
    strip.text = element_text(color = "#afafaf", size = 7.4),
    # axis title
    axis.title = element_text(color = "#666666", size = 8),
    # axis text
    axis.text.x = element_text(color = "#afafaf", size = 7.4, vjust = 0.9),
    axis.text.y = element_text(color = "darkgray", size = 7),
    # axis ticks
    axis.ticks.x = element_line(colour = "#bdbdbd", size = 0.4),
    axis.ticks.y = element_blank(),
    # title
    plot.title = element_text(color = "#666666", size = 9.1, hjust = 0.5)
  )
```

\newpage

# intersectionality

```{r}
dec_frames %>% 
  select(case, decision_date, in_freq) %>% 
  mutate(year = lubridate::year(decision_date)) %>% 
  ggplot(., aes(x = reorder(case, year), y = in_freq)) +
  geom_bar(stat = "identity") +
  labs(
    x = "Case",
    y = "Frequency of frame words/phrases in decision (divided by total word count)",
    title = "Word/Phrase Count for Intersectionality Frame"
  ) + 
  facet_grid(~ year, switch = "x", scales = "free_x", space = "free_x") +
  theme_bw() +
  theme(
    # grid lines + background
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "white", size = 0.1),
    panel.border = element_blank(),
    plot.background = element_rect(color = "#eaeaea", size = 0.4),
    axis.line = element_line(color = "#eaeaea", size = 0.4),
    # facet format
    panel.spacing = unit(0, "pt"), 
    strip.background = element_blank(),
    strip.placement = "outside",
    strip.text = element_text(color = "#afafaf", size = 7.4),
    # axis title
    axis.title = element_text(color = "#666666", size = 8),
    # axis text
    axis.text.x = element_text(color = "#afafaf", size = 7.4, vjust = 0.9),
    axis.text.y = element_text(color = "darkgray", size = 7),
    # axis ticks
    axis.ticks.x = element_line(colour = "#bdbdbd", size = 0.4),
    axis.ticks.y = element_blank(),
    # title
    plot.title = element_text(color = "#666666", size = 9.1, hjust = 0.5)
  )
```

\newpage

# morality

```{r}
dec_frames %>% 
  select(case, decision_date, mo_freq) %>% 
  mutate(year = lubridate::year(decision_date)) %>% 
  ggplot(., aes(x = reorder(case, year), y = mo_freq)) +
  geom_bar(stat = "identity") +
  labs(
    x = "Case",
    y = "Frequency of frame words/phrases in decision (divided by total word count)",
    title = "Word/Phrase Count for Morality Frame"
  ) + 
  facet_grid(~ year, switch = "x", scales = "free_x", space = "free_x") +
  theme_bw() +
  theme(
    # grid lines + background
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "white", size = 0.1),
    panel.border = element_blank(),
    plot.background = element_rect(color = "#eaeaea", size = 0.4),
    axis.line = element_line(color = "#eaeaea", size = 0.4),
    # facet format
    panel.spacing = unit(0, "pt"), 
    strip.background = element_blank(),
    strip.placement = "outside",
    strip.text = element_text(color = "#afafaf", size = 7.4),
    # axis title
    axis.title = element_text(color = "#666666", size = 8),
    # axis text
    axis.text.x = element_text(color = "#afafaf", size = 7.4, vjust = 0.9),
    axis.text.y = element_text(color = "darkgray", size = 7),
    # axis ticks
    axis.ticks.x = element_line(colour = "#bdbdbd", size = 0.4),
    axis.ticks.y = element_blank(),
    # title
    plot.title = element_text(color = "#666666", size = 9.1, hjust = 0.5)
  )
```

\newpage

# psy_harm

```{r}
dec_frames %>% 
  select(case, decision_date, ph_freq) %>% 
  mutate(year = lubridate::year(decision_date)) %>% 
  ggplot(., aes(x = reorder(case, year), y = ph_freq)) +
  geom_bar(stat = "identity") +
  labs(
    x = "Case",
    y = "Frequency of frame words/phrases in decision (divided by total word count)",
    title = "Word/Phrase Count for Psycological Harm Frame"
  ) + 
  facet_grid(~ year, switch = "x", scales = "free_x", space = "free_x") +
  theme_bw() +
  theme(
    # grid lines + background
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "white", size = 0.1),
    panel.border = element_blank(),
    plot.background = element_rect(color = "#eaeaea", size = 0.4),
    axis.line = element_line(color = "#eaeaea", size = 0.4),
    # facet format
    panel.spacing = unit(0, "pt"), 
    strip.background = element_blank(),
    strip.placement = "outside",
    strip.text = element_text(color = "#afafaf", size = 7.4),
    # axis title
    axis.title = element_text(color = "#666666", size = 8),
    # axis text
    axis.text.x = element_text(color = "#afafaf", size = 7.4, vjust = 0.9),
    axis.text.y = element_text(color = "darkgray", size = 7),
    # axis ticks
    axis.ticks.x = element_line(colour = "#bdbdbd", size = 0.4),
    axis.ticks.y = element_blank(),
    # title
    plot.title = element_text(color = "#666666", size = 9.1, hjust = 0.5)
  )
```

\newpage

# womanCitizenship

```{r}
dec_frames %>% 
  select(case, decision_date, wc_freq) %>% 
  mutate(year = lubridate::year(decision_date)) %>% 
  ggplot(., aes(x = reorder(case, year), y = wc_freq)) +
  geom_bar(stat = "identity") +
  labs(
    x = "Case",
    y = "Frequency of frame words/phrases in decision (divided by total word count)",
    title = "Word/Phrase Count for Woman Citizenship Frame"
  ) + 
  facet_grid(~ year, switch = "x", scales = "free_x", space = "free_x") +
  theme_bw() +
  theme(
    # grid lines + background
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "white", size = 0.1),
    panel.border = element_blank(),
    plot.background = element_rect(color = "#eaeaea", size = 0.4),
    axis.line = element_line(color = "#eaeaea", size = 0.4),
    # facet format
    panel.spacing = unit(0, "pt"), 
    strip.background = element_blank(),
    strip.placement = "outside",
    strip.text = element_text(color = "#afafaf", size = 7.4),
    # axis title
    axis.title = element_text(color = "#666666", size = 8),
    # axis text
    axis.text.x = element_text(color = "#afafaf", size = 7.4, vjust = 0.9),
    axis.text.y = element_text(color = "darkgray", size = 7),
    # axis ticks
    axis.ticks.x = element_line(colour = "#bdbdbd", size = 0.4),
    axis.ticks.y = element_blank(),
    # title
    plot.title = element_text(color = "#666666", size = 9.1, hjust = 0.5)
  )
```