### 2022-05-10 data-cleaning/00-scrape-amicus-text.Rmd
add litigant frames

---
title: "Tidy Amicus Data"
author: "Amanda Konet"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F, error = F)

# Box access
library(boxr)
box_auth()

# data manip?
library(tidyverse)

# parallel processing
library(R.utils)
library(doParallel)
```

# Purpose

There are ~40 folders in Amicus IV > data > amicus files. For each amicus file (denoted by files with "amicus" or "amici" in the file name), copy the file ID, file name, case name, and file text into a CSV.

# 1. Obtain file names and IDs

Set working directory to amcius files folder

```{r}
box_setwd("146066009084")
```

Get IDs of all the case folders

```{r}
# Get the IDs of all the case folders 
case_folders <- box_ls() %>%
  as.data.frame() %>% 
  filter(type == "folder") %>% 
  rename(case = name)

case_folders %>% head(1)
```

If only getting specific folder, change here

```{r}
#case_folders <- case_folders %>% filter(case == "Gonzales v PP")
```

For all of the folders, retrieve every file name and ID.

```{r}
# number of cores available
n_cores <- detectCores() - 1
registerDoParallel(n_cores)

# given a df of ids for SC folder, extract names and ids of all files in case folder
# then, save only the name and id of the decision file
all_file_names <- function(case_folder_ids) {

  # get info for all files contained in each case folder
  # stored in a list of lists
  cases_ids_list <- mclapply(case_folder_ids$id,
                                box_ls,
                                mc.cores = n_cores)

  # convert to list of df
  cases_ids_list <- map(cases_ids_list, as.data.frame)

  # get name of folder (this is case name)
  # NOTE path col is Amicus IV/data/amicus files/case name - we want the last bit "case name"
  cases_ids_list <- lapply(cases_ids_list, function(x) transform(x, case = sub(".*/", "", path)))

  print("Created list of dfs; each element is a case folder")

  # create iterator that will go through list element by element
  it <- iter(cases_ids_list, by = "row")

  # in each case folder, grab the name and id of the decision file
  file_ids <-
    foreach(i = it,
            .combine = rbind,
            .inorder = F) %dopar% {
              i %>%
                filter(grepl("amicus|amici", name, ignore.case = T)) %>%
                select(case, name, id)
            }
  print("Obtained amicus file ids")

  return(file_ids)
}

# run the function for Abortion and Sex Discrim folders
amicus_file_ids <- all_file_names(case_folders)

doParallel::stopImplicitCluster()
```

Check what we have - count of amicus briefs by case for manual comparison

```{r}
# save this file to data > processed amicus files
amicus_file_ids %>% 
  right_join(., case_folders, "case") %>% 
  group_by(case) %>% 
  summarize(n = sum(grepl("amicus|amici", name, ignore.case = T))) %>% 
  arrange(case) #%>% 
  #box_write(., "amicus brief counts.xlsx", dir_id = "145990409196")
```

Save list of amicus file IDs on Box

```{r}
amicus_file_ids %>% 
  right_join(., case_folders, "case") %>% 
  select(-id.y) %>% 
  rename(id = id.x) %>% 
  mutate(file_type = tolower(sub(".*\\.", "", name))) %>% 
  box_write(., "amicus-file-ids.csv", dir_id = "145990409196")
```

# 2. Scrape Text

Obtain the raw text from each amicus brief. Note that we have two file types here: DOCX & PDF.

Add file extension as col

```{r}
amicus_file_ids <- amicus_file_ids %>% 
  mutate(file_type = tolower(sub(".*\\.", "", name))) 
amicus_file_ids %>% group_by(file_type) %>% summarize(n = n())
```
Scrape text

```{r}
# create iterator 
itd <- iter(amicus_file_ids, by = "row")

# set up cluster
n_cores <- detectCores() - 1
registerDoParallel(n_cores)

# get text of cases in parallel
amicus_txt <- foreach(i = itd, .inorder = F, .combine = rbind) %dopar% {

  # get id and file type of current file
  id <- i %>% pull(id)
  type <- i %>% pull(file_type)

  # grab case and brief names
  case <- i %>% pull(case)
  brief <- i %>% pull(name)

  if (type == "pdf") {
    # is the file a PDF

    txt <- box_read(
      file_id = id,
      read_fun = function(x)
        pdftools::pdf_text(x) %>%
        readr::read_lines()
    )

  } else if (type == "docx") {
    # is the file a DOCX

    txt <- box_read(
      file_id = id,
      read_fun = function(x)
        textreadr::read_docx(x)
    )

  }  else {
    txt <- box_read(
      file_id = id,
      read_fun = function(x)
        textreadr::read_document(x)
    )

  }

  # grab text and concat to string
  #txt <- box_read(
  #  file_id = id,
  #  read_fun = function(x)
  #    textreadr::read_document(x)
  #)

  # concatenate to string
  txt_full <- paste(txt, sep = ' ', collapse = ' ')

  # add row
  data.frame(case, brief, id, txt_full)
}

doParallel::stopImplicitCluster()
```

Check that our files contain "summary of (the) argument: this is the phrase that indicates the start of the text we want to analyze.

```{r}
amicus_txt %>% 
  mutate(txt_full = str_squish(txt_full),
         summary = grepl("summary of argument|summary of the argument", txt_full, ignore.case = T)) %>%
  group_by(summary) %>% 
  summarize(n=n())
```

Which files do not? -- should just be the 3 that we have manual adjustments recorded for.

```{r}
amicus_txt %>% 
  mutate(txt_full = str_squish(txt_full),
         summary = grepl("summary of argument|summary of the argument", txt_full, ignore.case = T)) %>%
  filter(!summary) %>% 
  select(brief)
```

Final column names

```{r}
amicus_txt %>% colnames()
```

If adding in briefs, run below

```{r}
#full <- box_read("1029611016764")
#full <- full %>% rbind(amicus_txt) %>% arrange(case)
```

# 3. Write file to Box

```{r}
box_write(amicus_txt,
          file_name = "raw-amicus-brief-text.csv",
          dir_id = '145990409196')
```