### 2022-05-15 modeling/justicemq-interactions/legauf.Rmd

---
title: "legauf Amicus Brief Models"
author: "Amanda Konet"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, error = F, message = F)

library(tidyverse)
library(lme4)
library(kableExtra)
library(broom.mixed)
library(ggplot2)
library(effects)
library(interactions)
library(boxr)
box_auth()

amicus_only <- "948366662469"
full_cases <- "948832669583"
final_cases <- "949901868853"
```

# Data

Filter out instances where we don't have the dependent variable, justicevot. Change data types where necessary. Fill in NAs for frame freq with 0s

```{r}
data <- box_read(final_cases) %>% 
  filter(justicevot != "NA") %>% 
  mutate(justicevot = as.numeric(justicevot),
         justiceid = as.factor(justiceid),
         justicefem = as.factor(justicefem),
         nonftyp2 = as.factor(nonftyp2),
         lcdisdirfem = ifelse(lcdisdirfem==1, 0, 1),
         lcdisdirfem = as.factor(lcdisdirfem),
         #justicemq = cut(justicemq, breaks = c(-8, -2, 2, 5), labels = c("lib", "mod", "consv"),
         #               include.lowest = T, right = F),
         #justicemq = ifelse(justicemq >= 0, 0, 1),
         #justicemq = as.factor(justicemq),
         cirsplit = as.factor(cirsplit)) %>%
  mutate(across(c(legauf_abortion_different:legauf_womens_rights), ~replace_na(.x, 0))) %>% 
  filter(case != "Gonzales v PP")

justicemq_conv <- data %>% select(orig = justicemq) %>% 
  mutate(scaled = scale(orig)) %>%
  arrange(scaled)
```

Add Case ID-- this will be the second level var

```{r}
case_ids <- data %>% distinct(case) %>% arrange(case) %>% mutate(id = 1:n())
data <- case_ids %>% left_join(., data, by = "case")
#data %>% head(1) %>% select(case:justicefem)
```

# Scale Data

Remove cols where all values are NA. Not sure if we will do this or not in models.

```{r}
data <- data %>%
  select(case:cirsplit, pfemam, legauf_abortion_different:legauf_womens_rights) %>% 
  mutate(across(c(where(is.numeric), -justicevot, -id, -cirsplit),
                ~(scale(., center=T, scale=T))))
                #~(scale(., center=T, scale=T) %>% as.vector)))

total_cols <- data %>% select(contains('legauf')) %>% colnames()
data <- janitor::remove_empty(data, which = "cols")
remaining_cols <- data %>% select(contains('legauf')) %>% colnames()
#setdiff(total_cols, remaining_cols)
```

```{r, eval=F}
# Full List of Frames

_abortion_different +
_abortion_immoral +
_anti_women +
_bad_law + 
_bodily_integrity + 
_compulsory_motherhood +   
_compulsory_motherhoood_ii +
_couple +
_equal_protection +
_evidence  +
_favor_childbirth +
_fetus +
_final_authority +
_informed_consent +
_intersectionality +
_morality_religion +
_overturn_roe  +
_physician +
_physician_negative +
_privacy  +
_protesters +               
_standing +
_state_interest +
_uncertainty +
_undue_burden +
_untruthful +
_womens_future_life + 
_womens_health + 
womens_rights
```

# Significance

The following frames are significant: (see excel)

* 0:
* 0.01
* 0.05: 
* 0.10: 
* 0.20:

\newpage

# Abortion Regret

```{r}
mod3 <- glmer(justicevot ~
                (1|id) + justicemq + justicefem + lcdisdirfem + pfemam + 
                legauf_abortion_regret + legauf_abortion_regret:justicemq,
              family = binomial, data = data,  
              nAGQ = 0, control=glmerControl(
                check.conv.grad="ignore",check.conv.singular="ignore",check.conv.hess="ignore"))

tidy(mod3,conf.int=F,exponentiate=T,effects="fixed") %>% 
  mutate(across(where(is.numeric), ~round(.x, 4))) %>% 
  mutate(probability_yes = round(estimate/(1+estimate),4)) %>% 
  mutate(confidence = case_when(
    p.value >= 0 & p.value <= 0.001 ~ "***", 
    p.value > 0.001 & p.value <= 0.01 ~ "**",
    p.value > 0.01 & p.value <= 0.05 ~ "*",
    p.value > 0.05 & p.value <= 0.1 ~ ".",
    T ~ ""
  )) %>% 
  rename(odds_ratio = estimate) %>% 
  select(term, odds_ratio, p_value = p.value, confidence) %>% 
  kable()

```

<br>

```{r}
interact_plot(mod3, pred = legauf_abortion_regret, modx = justicemq,
              #modx.values = c(-3.5, -1, 0, 1, 3.5),
              # random
              #modx.values = c(-1.65, -.55, -.08, .35, 1.5),
              # quantiles
              modx.values = c(-3.4, -.75, -.3, .59, 1.9),
              modx.labels = c("More Liberal", "Moderately Liberal","Moderate", "Moderately Conservative", "More Conservative"),
              colors = c("blue", "gray", "lightskyblue", "darksalmon", "red"),
              #outcome.scale = "link",
              #interval = T,
              x.label = "Frame Frequency",
              y.label = "Probability of Voting in Favor of Repro. Rights")
```

```{r, eval=F}
# interaction levels
ef <- effect(term="justicemq:legauf_abortion_regret", mod = mod3)
#ef1<-effect(term="pfemau*justicemq",xlevels=list(pfemau=c(0,.5,1)),mod=res0)  #used this in paper

# as dataframe
ef_df <-as.data.frame(ef) %>% mutate(legauf_abortion_regret = as.factor(legauf_abortion_regret))

#By Color 
ggplot(ef_df, aes(x = justicemq, y= fit, group = legauf_abortion_regret))+
  geom_point()+
  geom_line(size = 1, aes(linetype = legauf_abortion_regret))+
  geom_ribbon(aes(ymin=fit-se, ymax=fit+se, fill=legauf_abortion_regret),alpha=0.5) + 
  labs(title = "Frequency of Abortion Regret Frame and Justice Ideology", 
       x= "Liberal                        Justice Ideology                         Conservative",
       y="Vote for Opp.             Marginal Effects             Vote for Fem.") + 
  guides(fill=guide_legend(reverse=T))+
  guides(linetype=guide_legend(reverse=T))+
  theme_bw() +
  #scale_fill_grey()+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(fill="Frame Frequency",
       linetype="Frame Frequency")
```

\newpage

```{r}
summary(mod3)
```

\newpage

# Bad Law

```{r}
mod4 <- glmer(justicevot ~
                (1|id) + justicemq + justicefem + lcdisdirfem + pfemam + 
                legauf_bad_law + legauf_bad_law:justicemq,
              family = binomial, data = data,  
              nAGQ = 0, control=glmerControl(
                check.conv.grad="ignore",check.conv.singular="ignore",check.conv.hess="ignore"))

tidy(mod4,conf.int=F,exponentiate=T,effects="fixed") %>% 
  mutate(across(where(is.numeric), ~round(.x, 4))) %>% 
  mutate(probability_yes = round(estimate/(1+estimate),4)) %>% 
  mutate(confidence = case_when(
    p.value >= 0 & p.value <= 0.001 ~ "***", 
    p.value > 0.001 & p.value <= 0.01 ~ "**",
    p.value > 0.01 & p.value <= 0.05 ~ "*",
    p.value > 0.05 & p.value <= 0.1 ~ ".",
    T ~ ""
  )) %>% 
  rename(odds_ratio = estimate) %>% 
  select(term, odds_ratio, p_value = p.value, confidence) %>% 
  kable()

```

<br>

```{r}
interact_plot(mod4, pred = legauf_bad_law, modx = justicemq,
              #modx.values = c(-3.5, -1, 0, 1, 3.5),
              # random
              #modx.values = c(-1.65, -.55, -.08, .35, 1.5),
              # quantiles
              modx.values = c(-3.4, -.75, -.3, .59, 1.9),
              modx.labels = c("More Liberal", "Moderately Liberal","Moderate", "Moderately Conservative", "More Conservative"),
              colors = c("blue", "gray", "lightskyblue", "darksalmon", "red"),
              #outcome.scale = "link",
              x.label = "Frame Frequency",
              y.label = "Probability of Voting in Favor of Repro. Rights")
```

```{r, eval=F}
# interaction levels
ef <- effect(term="justicemq:legauf_bad_law", mod = mod4)
ef <- effect(term="justicemq:legauf_bad_law", xlevels = list(justicemq=c(-4,-1,1,4)) ,mod = mod4)  #used this in paper

# as dataframe
#ef_df <-as.data.frame(ef) %>% mutate(legauf_bad_law = as.factor(legauf_bad_law))
ef_df <-as.data.frame(ef) %>% mutate(justicemq = as.factor(justicemq))

#By Color 
ggplot(ef_df, aes(x = legauf_bad_law, y= fit, group = justicemq))+
  geom_point()+
  geom_line(size = 1, aes(linetype = justicemq))+
  geom_ribbon(aes(ymin=fit-se, ymax=fit+se, fill=justicemq),alpha=0.5) + 
  labs(title = "Frequency of Bad Law Frame and Justice Ideology", 
       x= "Liberal                        Justice Ideology                         Conservative",
       y="Vote for Opp.             Marginal Effects             Vote for Fem.") + 
  guides(fill=guide_legend(reverse=T))+
  guides(linetype=guide_legend(reverse=T))+
  theme_bw() +
  #scale_fill_grey()+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(fill="Frame Frequency",
       linetype="Frame Frequency")
```

\newpage

```{r}
summary(mod4)
```

\newpage

# Evidence

```{r}
mod9 <- glmer(justicevot ~
                (1|id) + justicemq + justicefem + lcdisdirfem + pfemam + 
                legauf_evidence + legauf_evidence:justicemq,
              family = binomial, data = data,  
              nAGQ = 0, control=glmerControl(
                check.conv.grad="ignore",check.conv.singular="ignore",check.conv.hess="ignore"))

tidy(mod9,conf.int=F,exponentiate=T,effects="fixed") %>% 
  mutate(across(where(is.numeric), ~round(.x, 4))) %>% 
  mutate(probability_yes = round(estimate/(1+estimate),4)) %>% 
  mutate(confidence = case_when(
    p.value >= 0 & p.value <= 0.001 ~ "***", 
    p.value > 0.001 & p.value <= 0.01 ~ "**",
    p.value > 0.01 & p.value <= 0.05 ~ "*",
    p.value > 0.05 & p.value <= 0.1 ~ ".",
    T ~ ""
  )) %>% 
  rename(odds_ratio = estimate) %>% 
  select(term, odds_ratio, p_value = p.value, confidence) %>% 
  kable()
```

<br>

```{r}
interact_plot(mod9, pred = legauf_evidence, modx = justicemq,
              #modx.values = c(-3.5, -1, 0, 1, 3.5),
              # random
              #modx.values = c(-1.65, -.55, -.08, .35, 1.5),
              # quantiles
              modx.values = c(-3.4, -.75, -.3, .59, 1.9),
              modx.labels = c("More Liberal", "Moderately Liberal","Moderate", "Moderately Conservative", "More Conservative"),
              colors = c("blue", "gray", "lightskyblue", "darksalmon", "red"),
              #outcome.scale = "link",
              x.label = "Frame Frequency",
              y.label = "Probability of Voting in Favor of Repro. Rights")
```

```{r, eval=F}
# interaction levels
ef <- effect(term="justicemq:legauf_evidence", mod = mod9)
#ef1<-effect(term="pfemau*justicemq",xlevels=list(pfemau=c(0,.5,1)),mod=res0)  #used this in paper

# as dataframe
ef_df <-as.data.frame(ef) %>% mutate(legauf_evidence = as.factor(legauf_evidence))

#By Color 
ggplot(ef_df, aes(x = justicemq, y= fit, group = legauf_evidence))+
  geom_point()+
  geom_line(size = 1, aes(linetype = legauf_evidence))+
  geom_ribbon(aes(ymin=fit-se, ymax=fit+se, fill=legauf_evidence),alpha=0.5) + 
  labs(title = "Frequency of Evidence Frame and Justice Ideology", 
       x= "Liberal                        Justice Ideology                         Conservative",
       y="Vote for Opp.             Marginal Effects             Vote for Fem.") + 
  guides(fill=guide_legend(reverse=T))+
  guides(linetype=guide_legend(reverse=T))+
  theme_bw() +
  #scale_fill_grey()+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(fill="Frame Frequency",
       linetype="Frame Frequency")
```

\newpage

```{r}
summary(mod9)
```

\newpage

# Fetus

```{r}
mod11 <- glmer(justicevot ~
                (1|id) + justicemq + justicefem + lcdisdirfem + pfemam + 
                legauf_fetus + legauf_fetus:justicemq,
              family = binomial, data = data,  
              nAGQ = 0, control=glmerControl(
                check.conv.grad="ignore",check.conv.singular="ignore",check.conv.hess="ignore"))

tidy(mod11,conf.int=F,exponentiate=T,effects="fixed") %>% 
  mutate(across(where(is.numeric), ~round(.x, 4))) %>% 
  mutate(probability_yes = round(estimate/(1+estimate),4)) %>% 
  mutate(confidence = case_when(
    p.value >= 0 & p.value <= 0.001 ~ "***", 
    p.value > 0.001 & p.value <= 0.01 ~ "**",
    p.value > 0.01 & p.value <= 0.05 ~ "*",
    p.value > 0.05 & p.value <= 0.1 ~ ".",
    T ~ ""
  )) %>% 
  rename(odds_ratio = estimate) %>% 
  select(term, odds_ratio, p_value = p.value, confidence) %>% 
  kable()

```

<br>

```{r}
interact_plot(mod11, pred = legauf_fetus, modx = justicemq,
              #modx.values = c(-3.5, -1, 0, 1, 3.5),
              # random
              #modx.values = c(-1.65, -.55, -.08, .35, 1.5),
              # quantiles
              modx.values = c(-3.4, -.75, -.3, .59, 1.9),
              modx.labels = c("More Liberal", "Moderately Liberal","Moderate", "Moderately Conservative", "More Conservative"),
              colors = c("blue", "gray", "lightskyblue", "darksalmon", "red"),
              #outcome.scale = "link",
              x.label = "Frame Frequency",
              y.label = "Probability of Voting in Favor of Repro. Rights")
```

```{r, eval=F}
# interaction levels
ef <- effect(term="justicemq:legauf_fetus", mod = mod11)
#ef1<-effect(term="pfemau*justicemq",xlevels=list(pfemau=c(0,.5,1)),mod=res0)  #used this in paper

# as dataframe
ef_df <-as.data.frame(ef) %>% mutate(legauf_fetus = as.factor(legauf_fetus))

#By Color 
ggplot(ef_df, aes(x = justicemq, y= fit, group = legauf_fetus))+
  geom_point()+
  geom_line(size = 1, aes(linetype = legauf_fetus))+
  geom_ribbon(aes(ymin=fit-se, ymax=fit+se, fill=legauf_fetus),alpha=0.5) + 
  labs(title = "Frequency of Fetus Frame and Justice Ideology", 
       x= "Liberal                        Justice Ideology                         Conservative",
       y="Vote for Opp.             Marginal Effects             Vote for Fem.") + 
  guides(fill=guide_legend(reverse=T))+
  guides(linetype=guide_legend(reverse=T))+
  theme_bw() +
  #scale_fill_grey()+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(fill="Frame Frequency",
       linetype="Frame Frequency")
```

\newpage

```{r}
summary(mod11)
```

\newpage

# Final Authority

```{r}
mod12 <- glmer(justicevot ~
                (1|id) + justicemq + justicefem + lcdisdirfem + pfemam + 
                legauf_final_authority + legauf_final_authority:justicemq,
              family = binomial, data = data,  
              nAGQ = 0, control=glmerControl(
                check.conv.grad="ignore",check.conv.singular="ignore",check.conv.hess="ignore"))

tidy(mod12,conf.int=F,exponentiate=T,effects="fixed") %>% 
  mutate(across(where(is.numeric), ~round(.x, 4))) %>% 
  mutate(probability_yes = round(estimate/(1+estimate),4)) %>% 
  mutate(confidence = case_when(
    p.value >= 0 & p.value <= 0.001 ~ "***", 
    p.value > 0.001 & p.value <= 0.01 ~ "**",
    p.value > 0.01 & p.value <= 0.05 ~ "*",
    p.value > 0.05 & p.value <= 0.1 ~ ".",
    T ~ ""
  )) %>% 
  rename(odds_ratio = estimate) %>% 
  select(term, odds_ratio, p_value = p.value, confidence) %>% 
  kable()

```

<br>

```{r}
interact_plot(mod12, pred = legauf_final_authority, modx = justicemq,
              #modx.values = c(-3.5, -1, 0, 1, 3.5),
              # random
              #modx.values = c(-1.65, -.55, -.08, .35, 1.5),
              # quantiles
              modx.values = c(-3.4, -.75, -.3, .59, 1.9),
              modx.labels = c("More Liberal", "Moderately Liberal","Moderate", "Moderately Conservative", "More Conservative"),
              colors = c("blue", "gray", "lightskyblue", "darksalmon", "red"),
              #outcome.scale = "link",
              x.label = "Frame Frequency",
              y.label = "Probability of Voting in Favor of Repro. Rights")
```

```{r, eval=F}
# interaction levels
ef <- effect(term="justicemq:legauf_final_authority", mod = mod12)
#ef1<-effect(term="pfemau*justicemq",xlevels=list(pfemau=c(0,.5,1)),mod=res0)  #used this in paper

# as dataframe
ef_df <-as.data.frame(ef) %>% mutate(legauf_final_authority = as.factor(legauf_final_authority))

#By Color 
ggplot(ef_df, aes(x = justicemq, y= fit, group = legauf_final_authority))+
  geom_point()+
  geom_line(size = 1, aes(linetype = legauf_final_authority))+
  geom_ribbon(aes(ymin=fit-se, ymax=fit+se, fill=legauf_final_authority),alpha=0.5) + 
  labs(title = "Frequency of Final Authority Frame and Justice Ideology", 
       x= "Liberal                        Justice Ideology                         Conservative",
       y="Vote for Opp.             Marginal Effects             Vote for Fem.") + 
  guides(fill=guide_legend(reverse=T))+
  guides(linetype=guide_legend(reverse=T))+
  theme_bw() +
  #scale_fill_grey()+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(fill="Frame Frequency",
       linetype="Frame Frequency")
```

\newpage

```{r}
summary(mod12)
```
\newpage

# Morality Religion

```{r}
mod14 <- glmer(justicevot ~
                (1|id) + justicemq + justicefem + lcdisdirfem + pfemam + 
                legauf_morality_religion + legauf_morality_religion:justicemq,
              family = binomial, data = data,  
              nAGQ = 0, control=glmerControl(
                check.conv.grad="ignore",check.conv.singular="ignore",check.conv.hess="ignore"))

tidy(mod14,conf.int=F,exponentiate=T,effects="fixed") %>% 
  mutate(across(where(is.numeric), ~round(.x, 4))) %>% 
  mutate(probability_yes = round(estimate/(1+estimate),4)) %>% 
  mutate(confidence = case_when(
    p.value >= 0 & p.value <= 0.001 ~ "***", 
    p.value > 0.001 & p.value <= 0.01 ~ "**",
    p.value > 0.01 & p.value <= 0.05 ~ "*",
    p.value > 0.05 & p.value <= 0.1 ~ ".",
    T ~ ""
  )) %>% 
  rename(odds_ratio = estimate) %>% 
  select(term, odds_ratio, p_value = p.value, confidence) %>% 
  kable()

```

<br>

```{r}
interact_plot(mod14, pred = legauf_morality_religion, modx = justicemq,
              #modx.values = c(-3.5, -1, 0, 1, 3.5),
              # random
              #modx.values = c(-1.65, -.55, -.08, .35, 1.5),
              # quantiles
              modx.values = c(-3.4, -.75, -.3, .59, 1.9),
              modx.labels = c("More Liberal", "Moderately Liberal","Moderate", "Moderately Conservative", "More Conservative"),
              colors = c("blue", "gray", "lightskyblue", "darksalmon", "red"),
              #outcome.scale = "link",
              x.label = "Frame Frequency",
              y.label = "Probability of Voting in Favor of Repro. Rights")
```

```{r, eval=F}
# interaction levels
ef <- effect(term="justicemq:legauf_morality_religion", mod = mod14)
#ef1<-effect(term="pfemau*justicemq",xlevels=list(pfemau=c(0,.5,1)),mod=res0)  #used this in paper

# as dataframe
ef_df <-as.data.frame(ef) %>% mutate(legauf_morality_religion = as.factor(legauf_morality_religion))

#By Color 
ggplot(ef_df, aes(x = justicemq, y= fit, group = legauf_morality_religion))+
  geom_point()+
  geom_line(size = 1, aes(linetype = legauf_morality_religion))+
  geom_ribbon(aes(ymin=fit-se, ymax=fit+se, fill=legauf_morality_religion),alpha=0.5) + 
  labs(title = "Frequency of Morality & Religion Frame and Justice Ideology", 
       x= "Liberal                        Justice Ideology                         Conservative",
       y="Vote for Opp.             Marginal Effects             Vote for Fem.") + 
  guides(fill=guide_legend(reverse=T))+
  guides(linetype=guide_legend(reverse=T))+
  theme_bw() +
  #scale_fill_grey()+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(fill="Frame Frequency",
       linetype="Frame Frequency")
```

\newpage

```{r}
summary(mod14)
```

\newpage

# Overturn Roe

```{r}
mod15 <- glmer(justicevot ~
                (1|id) + justicemq + justicefem + lcdisdirfem + pfemam + 
                legauf_overturn_roe + legauf_overturn_roe:justicemq,
              family = binomial, data = data,  
              nAGQ = 0, control=glmerControl(
                check.conv.grad="ignore",check.conv.singular="ignore",check.conv.hess="ignore"))

tidy(mod15,conf.int=F,exponentiate=T,effects="fixed") %>% 
  mutate(across(where(is.numeric), ~round(.x, 4))) %>% 
  mutate(probability_yes = round(estimate/(1+estimate),4)) %>% 
  mutate(confidence = case_when(
    p.value >= 0 & p.value <= 0.001 ~ "***", 
    p.value > 0.001 & p.value <= 0.01 ~ "**",
    p.value > 0.01 & p.value <= 0.05 ~ "*",
    p.value > 0.05 & p.value <= 0.1 ~ ".",
    T ~ ""
  )) %>% 
  rename(odds_ratio = estimate) %>% 
  select(term, odds_ratio, p_value = p.value, confidence) %>% 
  kable()

```

<br>

```{r}
interact_plot(mod15, pred = legauf_overturn_roe, modx = justicemq,
              #modx.values = c(-3.5, -1, 0, 1, 3.5),
              # random
              #modx.values = c(-1.65, -.55, -.08, .35, 1.5),
              # quantiles
              modx.values = c(-3.4, -.75, -.3, .59, 1.9),
              modx.labels = c("More Liberal", "Moderately Liberal","Moderate", "Moderately Conservative", "More Conservative"),
              colors = c("blue", "gray", "lightskyblue", "darksalmon", "red"),
              #outcome.scale = "link",
              x.label = "Frame Frequency",
              y.label = "Probability of Voting in Favor of Repro. Rights")
```

```{r, eval=F}
# interaction levels
ef <- effect(term="justicemq:legauf_overturn_roe", mod = mod15)
#ef1<-effect(term="pfemau*justicemq",xlevels=list(pfemau=c(0,.5,1)),mod=res0)  #used this in paper

# as dataframe
ef_df <-as.data.frame(ef) %>% mutate(legauf_overturn_roe = as.factor(legauf_overturn_roe))

#By Color 
ggplot(ef_df, aes(x = justicemq, y= fit, group = legauf_overturn_roe))+
  geom_point()+
  geom_line(size = 1, aes(linetype = legauf_overturn_roe))+
  geom_ribbon(aes(ymin=fit-se, ymax=fit+se, fill=legauf_overturn_roe),alpha=0.5) + 
  labs(title = "Frequency of Overturn Roe Frame and Justice Ideology", 
       x= "Liberal                        Justice Ideology                         Conservative",
       y="Vote for Opp.             Marginal Effects             Vote for Fem.") + 
  guides(fill=guide_legend(reverse=T))+
  guides(linetype=guide_legend(reverse=T))+
  theme_bw() +
  #scale_fill_grey()+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(fill="Frame Frequency",
       linetype="Frame Frequency")
```

\newpage

```{r}
summary(mod15)
```

\newpage

# Privacy

```{r}
mod18 <- glmer(justicevot ~
               (1|id) + justicemq + justicefem + lcdisdirfem + pfemam + 
                legauf_privacy + legauf_privacy:justicemq,
              family = binomial, data = data,  
              nAGQ = 0, control=glmerControl(
                check.conv.grad="ignore",check.conv.singular="ignore",check.conv.hess="ignore"))

tidy(mod18,conf.int=F,exponentiate=T,effects="fixed") %>% 
  mutate(across(where(is.numeric), ~round(.x, 4))) %>% 
  mutate(probability_yes = round(estimate/(1+estimate),4)) %>% 
  mutate(confidence = case_when(
    p.value >= 0 & p.value <= 0.001 ~ "***", 
    p.value > 0.001 & p.value <= 0.01 ~ "**",
    p.value > 0.01 & p.value <= 0.05 ~ "*",
    p.value > 0.05 & p.value <= 0.1 ~ ".",
    T ~ ""
  )) %>% 
  rename(odds_ratio = estimate) %>% 
  select(term, odds_ratio, p_value = p.value, confidence) %>% 
  kable()

```

<br>

```{r}
interact_plot(mod18, pred = legauf_privacy, modx = justicemq,
              #modx.values = c(-3.5, -1, 0, 1, 3.5),
              # random
              #modx.values = c(-1.65, -.55, -.08, .35, 1.5),
              # quantiles
              modx.values = c(-3.4, -.75, -.3, .59, 1.9),
              modx.labels = c("More Liberal", "Moderately Liberal","Moderate", "Moderately Conservative", "More Conservative"),
              colors = c("blue", "gray", "lightskyblue", "darksalmon", "red"),
              #outcome.scale = "link",
              x.label = "Frame Frequency",
              y.label = "Probability of Voting in Favor of Repro. Rights")
```

```{r, eval=F}
# interaction levels
ef <- effect(term="justicemq:legauf_privacy", mod = mod18)
#ef1<-effect(term="pfemau*justicemq",xlevels=list(pfemau=c(0,.5,1)),mod=res0)  #used this in paper

# as dataframe
ef_df <-as.data.frame(ef) %>% mutate(legauf_privacy = as.factor(legauf_privacy))

#By Color 
ggplot(ef_df, aes(x = justicemq, y= fit, group = legauf_privacy))+
  geom_point()+
  geom_line(size = 1, aes(linetype = legauf_privacy))+
  geom_ribbon(aes(ymin=fit-se, ymax=fit+se, fill=legauf_privacy),alpha=0.5) + 
  labs(title = "Frequency of Privacy Frame and Justice Ideology", 
       x= "Liberal                        Justice Ideology                         Conservative",
       y="Vote for Opp.             Marginal Effects             Vote for Fem.") + 
  guides(fill=guide_legend(reverse=T))+
  guides(linetype=guide_legend(reverse=T))+
  theme_bw() +
  #scale_fill_grey()+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(fill="Frame Frequency",
       linetype="Frame Frequency")
```

\newpage

```{r}
summary(mod18)
```

\newpage

# Standing

```{r}
mod20 <- glmer(justicevot ~
                (1|id) + justicemq + justicefem + lcdisdirfem + pfemam + 
                legauf_standing + legauf_standing:justicemq,
              family = binomial, data = data,  
              nAGQ = 0, control=glmerControl(
                check.conv.grad="ignore",check.conv.singular="ignore",check.conv.hess="ignore"))

tidy(mod20,conf.int=F,exponentiate=T,effects="fixed") %>% 
  mutate(across(where(is.numeric), ~round(.x, 4))) %>% 
  mutate(probability_yes = round(estimate/(1+estimate),4)) %>% 
  mutate(confidence = case_when(
    p.value >= 0 & p.value <= 0.001 ~ "***", 
    p.value > 0.001 & p.value <= 0.01 ~ "**",
    p.value > 0.01 & p.value <= 0.05 ~ "*",
    p.value > 0.05 & p.value <= 0.1 ~ ".",
    T ~ ""
  )) %>% 
  rename(odds_ratio = estimate) %>% 
  select(term, odds_ratio, p_value = p.value, confidence) %>% 
  kable()

```

<br>

```{r}
interact_plot(mod20, pred = legauf_standing, modx = justicemq,
              #modx.values = c(-3.5, -1, 0, 1, 3.5),
              # random
              #modx.values = c(-1.65, -.55, -.08, .35, 1.5),
              # quantiles
              modx.values = c(-3.4, -.75, -.3, .59, 1.9),
              modx.labels = c("More Liberal", "Moderately Liberal","Moderate", "Moderately Conservative", "More Conservative"),
              colors = c("blue", "gray", "lightskyblue", "darksalmon", "red"),
              #outcome.scale = "link",
              x.label = "Frame Frequency",
              y.label = "Probability of Voting in Favor of Repro. Rights")
```

```{r, eval=F}
# interaction levels
ef <- effect(term="justicemq:legauf_standing", mod = mod20)
#ef1<-effect(term="pfemau*justicemq",xlevels=list(pfemau=c(0,.5,1)),mod=res0)  #used this in paper

# as dataframe
ef_df <-as.data.frame(ef) %>% mutate(legauf_standing = as.factor(legauf_standing))

#By Color 
ggplot(ef_df, aes(x = justicemq, y= fit, group = legauf_standing))+
  geom_point()+
  geom_line(size = 1, aes(linetype = legauf_standing))+
  geom_ribbon(aes(ymin=fit-se, ymax=fit+se, fill=legauf_standing),alpha=0.5) + 
  labs(title = "Frequency of Standing Frame and Justice Ideology", 
       x= "Liberal                        Justice Ideology                         Conservative",
       y="Vote for Opp.             Marginal Effects             Vote for Fem.") + 
  guides(fill=guide_legend(reverse=T))+
  guides(linetype=guide_legend(reverse=T))+
  theme_bw() +
  #scale_fill_grey()+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(fill="Frame Frequency",
       linetype="Frame Frequency")
```

\newpage

```{r}
summary(mod20)
```

\newpage

# Uncertainty

```{r}
mod22 <- glmer(justicevot ~
                (1|id) + justicemq + justicefem + lcdisdirfem + pfemam + 
                legauf_uncertainty + legauf_uncertainty:justicemq,
              family = binomial, data = data,  
              nAGQ = 0, control=glmerControl(
                check.conv.grad="ignore",check.conv.singular="ignore",check.conv.hess="ignore"))

tidy(mod22,conf.int=F,exponentiate=T,effects="fixed") %>% 
  mutate(across(where(is.numeric), ~round(.x, 4))) %>% 
  mutate(probability_yes = round(estimate/(1+estimate),4)) %>% 
  mutate(confidence = case_when(
    p.value >= 0 & p.value <= 0.001 ~ "***", 
    p.value > 0.001 & p.value <= 0.01 ~ "**",
    p.value > 0.01 & p.value <= 0.05 ~ "*",
    p.value > 0.05 & p.value <= 0.1 ~ ".",
    T ~ ""
  )) %>% 
  rename(odds_ratio = estimate) %>% 
  select(term, odds_ratio, p_value = p.value, confidence) %>% 
  kable()

```

<br>

```{r}
interact_plot(mod22, pred = legauf_uncertainty, modx = justicemq,
              #modx.values = c(-3.5, -1, 0, 1, 3.5),
              # random
              #modx.values = c(-1.65, -.55, -.08, .35, 1.5),
              # quantiles
              modx.values = c(-3.4, -.75, -.3, .59, 1.9),
              modx.labels = c("More Liberal", "Moderately Liberal","Moderate", "Moderately Conservative", "More Conservative"),
              colors = c("blue", "gray", "lightskyblue", "darksalmon", "red"),
              #outcome.scale = "link",
              x.label = "Frame Frequency",
              y.label = "Probability of Voting in Favor of Repro. Rights")
```

```{r, eval=F}
# interaction levels
ef <- effect(term="justicemq:legauf_uncertainty", mod = mod22)
#ef1<-effect(term="pfemau*justicemq",xlevels=list(pfemau=c(0,.5,1)),mod=res0)  #used this in paper

# as dataframe
ef_df <-as.data.frame(ef) %>% mutate(legauf_uncertainty = as.factor(legauf_uncertainty))

#By Color 
ggplot(ef_df, aes(x = justicemq, y= fit, group = legauf_uncertainty))+
  geom_point()+
  geom_line(size = 1, aes(linetype = legauf_uncertainty))+
  geom_ribbon(aes(ymin=fit-se, ymax=fit+se, fill=legauf_uncertainty),alpha=0.5) + 
  labs(title = "Frequency of Uncertainty Frame and Justice Ideology", 
       x= "Liberal                        Justice Ideology                         Conservative",
       y="Vote for Opp.             Marginal Effects             Vote for Fem.") + 
  guides(fill=guide_legend(reverse=T))+
  guides(linetype=guide_legend(reverse=T))+
  theme_bw() +
  #scale_fill_grey()+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(fill="Frame Frequency",
       linetype="Frame Frequency")
```

\newpage

```{r}
summary(mod22)
```

\newpage

# Undue Burden

```{r}
mod23 <- glmer(justicevot ~
                (1|id) + justicemq + justicefem + lcdisdirfem + pfemam + 
                legauf_undue_burden + legauf_undue_burden:justicemq,
              family = binomial, data = data,  
              nAGQ = 0, control=glmerControl(
                check.conv.grad="ignore",check.conv.singular="ignore",check.conv.hess="ignore"))

tidy(mod23,conf.int=F,exponentiate=T,effects="fixed") %>% 
  mutate(across(where(is.numeric), ~round(.x, 4))) %>% 
  mutate(probability_yes = round(estimate/(1+estimate),4)) %>% 
  mutate(confidence = case_when(
    p.value >= 0 & p.value <= 0.001 ~ "***", 
    p.value > 0.001 & p.value <= 0.01 ~ "**",
    p.value > 0.01 & p.value <= 0.05 ~ "*",
    p.value > 0.05 & p.value <= 0.1 ~ ".",
    T ~ ""
  )) %>% 
  rename(odds_ratio = estimate) %>% 
  select(term, odds_ratio, p_value = p.value, confidence) %>% 
  kable()

```

<br>

```{r}
interact_plot(mod23, pred = legauf_undue_burden, modx = justicemq,
              #modx.values = c(-3.5, -1, 0, 1, 3.5),
              # random
              #modx.values = c(-1.65, -.55, -.08, .35, 1.5),
              # quantiles
              modx.values = c(-3.4, -.75, -.3, .59, 1.9),
              modx.labels = c("More Liberal", "Moderately Liberal","Moderate", "Moderately Conservative", "More Conservative"),
              colors = c("blue", "gray", "lightskyblue", "darksalmon", "red"),
              #outcome.scale = "link",
              x.label = "Frame Frequency",
              y.label = "Probability of Voting in Favor of Repro. Rights")
```

```{r, eval=F}
# interaction levels
ef <- effect(term="justicemq:legauf_undue_burden", mod = mod23)
#ef1<-effect(term="pfemau*justicemq",xlevels=list(pfemau=c(0,.5,1)),mod=res0)  #used this in paper

# as dataframe
ef_df <-as.data.frame(ef) %>% mutate(legauf_undue_burden = as.factor(legauf_undue_burden))

#By Color 
ggplot(ef_df, aes(x = justicemq, y= fit, group = legauf_undue_burden))+
  geom_point()+
  geom_line(size = 1, aes(linetype = legauf_undue_burden))+
  geom_ribbon(aes(ymin=fit-se, ymax=fit+se, fill=legauf_undue_burden),alpha=0.5) + 
  labs(title = "Frequency of Undue Burden Frame and Justice Ideology", 
       x= "Liberal                        Justice Ideology                         Conservative",
       y="Vote for Opp.             Marginal Effects             Vote for Fem.") + 
  guides(fill=guide_legend(reverse=T))+
  guides(linetype=guide_legend(reverse=T))+
  theme_bw() +
  #scale_fill_grey()+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(fill="Frame Frequency",
       linetype="Frame Frequency")
```

\newpage

```{r}
summary(mod23)
```

\newpage

# Untruthful

```{r}
mod24 <- glmer(justicevot ~
                (1|id) + justicemq + justicefem + lcdisdirfem + pfemam + 
                legauf_untruthful + legauf_untruthful:justicemq,
              family = binomial, data = data,  
              nAGQ = 0, control=glmerControl(
                check.conv.grad="ignore",check.conv.singular="ignore",check.conv.hess="ignore"))

tidy(mod24,conf.int=F,exponentiate=T,effects="fixed") %>% 
  mutate(across(where(is.numeric), ~round(.x, 4))) %>% 
  mutate(probability_yes = round(estimate/(1+estimate),4)) %>% 
  mutate(confidence = case_when(
    p.value >= 0 & p.value <= 0.001 ~ "***", 
    p.value > 0.001 & p.value <= 0.01 ~ "**",
    p.value > 0.01 & p.value <= 0.05 ~ "*",
    p.value > 0.05 & p.value <= 0.1 ~ ".",
    T ~ ""
  )) %>% 
  rename(odds_ratio = estimate) %>% 
  select(term, odds_ratio, p_value = p.value, confidence) %>% 
  kable()

```

<br>

```{r}
interact_plot(mod24, pred = legauf_untruthful, modx = justicemq,
              #modx.values = c(-3.5, -1, 0, 1, 3.5),
              # random
              #modx.values = c(-1.65, -.55, -.08, .35, 1.5),
              # quantiles
              modx.values = c(-3.4, -.75, -.3, .59, 1.9),
              modx.labels = c("More Liberal", "Moderately Liberal","Moderate", "Moderately Conservative", "More Conservative"),
              colors = c("blue", "gray", "lightskyblue", "darksalmon", "red"),
              #outcome.scale = "link",
              x.label = "Frame Frequency",
              y.label = "Probability of Voting in Favor of Repro. Rights")
```

```{r, eval=F}
# interaction levels
ef <- effect(term="justicemq:legauf_untruthful", mod = mod24)
#ef1<-effect(term="pfemau*justicemq",xlevels=list(pfemau=c(0,.5,1)),mod=res0)  #used this in paper

# as dataframe
ef_df <-as.data.frame(ef) %>% mutate(legauf_untruthful = as.factor(legauf_untruthful))

#By Color 
ggplot(ef_df, aes(x = justicemq, y= fit, group = legauf_untruthful))+
  geom_point()+
  geom_line(size = 1, aes(linetype = legauf_untruthful))+
  geom_ribbon(aes(ymin=fit-se, ymax=fit+se, fill=legauf_untruthful),alpha=0.5) + 
  labs(title = "Frequency of Untruthful Frame and Justice Ideology", 
       x= "Liberal                        Justice Ideology                         Conservative",
       y="Vote for Opp.             Marginal Effects             Vote for Fem.") + 
  guides(fill=guide_legend(reverse=T))+
  guides(linetype=guide_legend(reverse=T))+
  theme_bw() +
  #scale_fill_grey()+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(fill="Frame Frequency",
       linetype="Frame Frequency")
```

\newpage

```{r}
summary(mod24)
```
\newpage

# Women's Future Life

```{r}
mod25 <- glmer(justicevot ~
                (1|id) + justicemq + justicefem + lcdisdirfem + pfemam + 
                legauf_womens_future_life + legauf_womens_future_life:justicemq,
              family = binomial, data = data,  
              nAGQ = 0, control=glmerControl(
                check.conv.grad="ignore",check.conv.singular="ignore",check.conv.hess="ignore"))

tidy(mod25,conf.int=F,exponentiate=T,effects="fixed") %>% 
  mutate(across(where(is.numeric), ~round(.x, 4))) %>% 
  mutate(probability_yes = round(estimate/(1+estimate),4)) %>% 
  mutate(confidence = case_when(
    p.value >= 0 & p.value <= 0.001 ~ "***", 
    p.value > 0.001 & p.value <= 0.01 ~ "**",
    p.value > 0.01 & p.value <= 0.05 ~ "*",
    p.value > 0.05 & p.value <= 0.1 ~ ".",
    T ~ ""
  )) %>% 
  rename(odds_ratio = estimate) %>% 
  select(term, odds_ratio, p_value = p.value, confidence) %>% 
  kable()

```

<br>

```{r}
interact_plot(mod25, pred = legauf_womens_future_life, modx = justicemq,
              #modx.values = c(-3.5, -1, 0, 1, 3.5),
              # random
              #modx.values = c(-1.65, -.55, -.08, .35, 1.5),
              # quantiles
              modx.values = c(-3.4, -.75, -.3, .59, 1.9),
              modx.labels = c("More Liberal", "Moderately Liberal","Moderate", "Moderately Conservative", "More Conservative"),
              colors = c("blue", "gray", "lightskyblue", "darksalmon", "red"),
              #outcome.scale = "link",
              x.label = "Frame Frequency",
              y.label = "Probability of Voting in Favor of Repro. Rights")
```

```{r, eval=F}
# interaction levels
ef <- effect(term="justicemq:legauf_womens_future_life", mod = mod25)
#ef1<-effect(term="pfemau*justicemq",xlevels=list(pfemau=c(0,.5,1)),mod=res0)  #used this in paper

# as dataframe
ef_df <-as.data.frame(ef) %>% mutate(legauf_womens_future_life = as.factor(legauf_womens_future_life))

#By Color 
ggplot(ef_df, aes(x = justicemq, y= fit, group = legauf_womens_future_life))+
  geom_point()+
  geom_line(size = 1, aes(linetype = legauf_womens_future_life))+
  geom_ribbon(aes(ymin=fit-se, ymax=fit+se, fill=legauf_womens_future_life),alpha=0.5) + 
  labs(title = "Frequency of Abortion Regret Frame and Justice Ideology", 
       x= "Liberal                        Justice Ideology                         Conservative",
       y="Vote for Opp.             Marginal Effects             Vote for Fem.") + 
  guides(fill=guide_legend(reverse=T))+
  guides(linetype=guide_legend(reverse=T))+
  theme_bw() +
  #scale_fill_grey()+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(fill="Frame Frequency",
       linetype="Frame Frequency")
```

\newpage

```{r}
summary(mod25)
```
\newpage

# Women's Health

```{r}
mod26 <- glmer(justicevot ~
                (1|id) + justicemq + justicefem + lcdisdirfem + pfemam + 
                legauf_womens_health + legauf_womens_health:justicemq,
              family = binomial, data = data,  
              nAGQ = 0, control=glmerControl(
                check.conv.grad="ignore",check.conv.singular="ignore",check.conv.hess="ignore"))

tidy(mod26,conf.int=F,exponentiate=T,effects="fixed") %>% 
  mutate(across(where(is.numeric), ~round(.x, 4))) %>% 
  mutate(probability_yes = round(estimate/(1+estimate),4)) %>% 
  mutate(confidence = case_when(
    p.value >= 0 & p.value <= 0.001 ~ "***", 
    p.value > 0.001 & p.value <= 0.01 ~ "**",
    p.value > 0.01 & p.value <= 0.05 ~ "*",
    p.value > 0.05 & p.value <= 0.1 ~ ".",
    T ~ ""
  )) %>% 
  rename(odds_ratio = estimate) %>% 
  select(term, odds_ratio, p_value = p.value, confidence) %>% 
  kable()

```

<br>

```{r}
interact_plot(mod26, pred = legauf_womens_health, modx = justicemq,
              #modx.values = c(-3.5, -1, 0, 1, 3.5),
              # random
              #modx.values = c(-1.65, -.55, -.08, .35, 1.5),
              # quantiles
              modx.values = c(-3.4, -.75, -.3, .59, 1.9),
              modx.labels = c("More Liberal", "Moderately Liberal","Moderate", "Moderately Conservative", "More Conservative"),
              colors = c("blue", "gray", "lightskyblue", "darksalmon", "red"),
              #outcome.scale = "link",
              x.label = "Frame Frequency",
              y.label = "Probability of Voting in Favor of Repro. Rights")
```

```{r, eval=F}
# interaction levels
ef <- effect(term="justicemq:legauf_womens_health", mod = mod26)
#ef1<-effect(term="pfemau*justicemq",xlevels=list(pfemau=c(0,.5,1)),mod=res0)  #used this in paper

# as dataframe
ef_df <-as.data.frame(ef) %>% mutate(legauf_womens_health = as.factor(legauf_womens_health))

#By Color 
ggplot(ef_df, aes(x = justicemq, y= fit, group = legauf_womens_health))+
  geom_point()+
  geom_line(size = 1, aes(linetype = legauf_womens_health))+
  geom_ribbon(aes(ymin=fit-se, ymax=fit+se, fill=legauf_womens_health),alpha=0.5) + 
  labs(title = "Frequency of Women's Health Frame and Justice Ideology", 
       x= "Liberal                        Justice Ideology                         Conservative",
       y="Vote for Opp.             Marginal Effects             Vote for Fem.") + 
  guides(fill=guide_legend(reverse=T))+
  guides(linetype=guide_legend(reverse=T))+
  theme_bw() +
  #scale_fill_grey()+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(fill="Frame Frequency",
       linetype="Frame Frequency")
```

\newpage

```{r}
summary(mod26)
```

\newpage

# Women's Rights

```{r}
mod27 <- glmer(justicevot ~
                (1|id) + justicemq + justicefem + lcdisdirfem + pfemam + 
                legauf_womens_rights + legauf_womens_rights:justicemq,
              family = binomial, data = data,  
              nAGQ = 0, control=glmerControl(
                check.conv.grad="ignore",check.conv.singular="ignore",check.conv.hess="ignore"))

tidy(mod27,conf.int=F,exponentiate=T,effects="fixed") %>% 
  mutate(across(where(is.numeric), ~round(.x, 4))) %>% 
  mutate(probability_yes = round(estimate/(1+estimate),4)) %>% 
  mutate(confidence = case_when(
    p.value >= 0 & p.value <= 0.001 ~ "***", 
    p.value > 0.001 & p.value <= 0.01 ~ "**",
    p.value > 0.01 & p.value <= 0.05 ~ "*",
    p.value > 0.05 & p.value <= 0.1 ~ ".",
    T ~ ""
  )) %>% 
  rename(odds_ratio = estimate) %>% 
  select(term, odds_ratio, p_value = p.value, confidence) %>% 
  kable()
```

<br>

```{r}
interact_plot(mod27, pred = legauf_womens_rights, modx = justicemq,
              #modx.values = c(-3.5, -1, 0, 1, 3.5),
              # random
              #modx.values = c(-1.65, -.55, -.08, .35, 1.5),
              # quantiles
              modx.values = c(-3.4, -.75, -.3, .59, 1.9),
              modx.labels = c("More Liberal", "Moderately Liberal","Moderate", "Moderately Conservative", "More Conservative"),
              colors = c("blue", "gray", "lightskyblue", "darksalmon", "red"),
              #outcome.scale = "link",
              x.label = "Frame Frequency",
              y.label = "Probability of Voting in Favor of Repro. Rights")
```

```{r, eval=F}
# interaction levels
ef <- effect(term="justicemq:legauf_womens_rights", mod = mod27)
#ef1<-effect(term="pfemau*justicemq",xlevels=list(pfemau=c(0,.5,1)),mod=res0)  #used this in paper

# as dataframe
ef_df <-as.data.frame(ef) %>% mutate(legauf_womens_rights = as.factor(legauf_womens_rights))

#By Color 
ggplot(ef_df, aes(x = justicemq, y= fit, group = legauf_womens_rights))+
  geom_point()+
  geom_line(size = 1, aes(linetype = legauf_womens_rights))+
  geom_ribbon(aes(ymin=fit-se, ymax=fit+se, fill=legauf_womens_rights),alpha=0.5) + 
  labs(title = "Frequency of Women's Rights Frame and Justice Ideology", 
       x= "Liberal                        Justice Ideology                         Conservative",
       y="Vote for Opp.             Marginal Effects             Vote for Fem.") + 
  guides(fill=guide_legend(reverse=T))+
  guides(linetype=guide_legend(reverse=T))+
  theme_bw() +
  #scale_fill_grey()+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(fill="Frame Frequency",
       linetype="Frame Frequency")
```

\newpage

```{r}
summary(mod27)
```