### 2022-05-24 eda/miscellaneous_eda.Rmd
eda for paper

---
title: "Miscellaneous"
author: "Sarah Torrence"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F, error = F)
library(tidyverse)
library(ggplot2)
library(boxr)
library(stargazer)
box_auth()

amicus <- box_read("947307207905", read_fun = readr::read_csv)
```

The average number of authors on the RR and AA amici (separately) for our 40 abortion cases

 (still in progress)

The average number of amici per case on the RR and AA sides (RR and AA separately).

```{r}
amici_per_case <- amicus %>% group_by(case, brief_party) %>% summarize(count = n())
amici_per_case %>% group_by(brief_party) %>% summarize(avg = mean(count))
```