### 2022-05-25 data-cleaning/07-generate-frame-measures.Rmd
update from when box wasnt working

---
title: "Generate Frame Measures"
author: "Amanda Konet"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F, error = F)
library(tidyverse)
library(tidytext)
library(textclean)
library(tm)
library(stringi)
library(quanteda)

library(boxr)
box_auth()
```

# Purpose

To develop amicus- and frame-level measures of frame usage

# Data

Abortion frame terms:

```{r}
frame_terms <- box_read("947335700912", read_fun = readr::read_csv)
#frame_terms <- read_csv("~/Box/Amicus IV/data/amicus contents - frames/clean-amicus-iv-frame-dictionary.csv")
```

Amicus data:

```{r}
# abortion cases
amicus <- box_read("947307207905", read_fun = readr::read_csv) %>% 
  mutate(brief_party = ifelse(id == 861815623586, 1, brief_party))

#amicus <- read_csv("~/Box/Amicus IV/data/processed amicus files/clean-shortened-amicus-text-vars.csv") %>% 
#  mutate(brief_party = ifelse(id == 861815623586, 1, brief_party))
```

# Amicus-Level Measures

Start by creating corpus

```{r}
amicus_corp <- amicus$text_clean %>% corpus()

# Add meta data to corpus
docvars(amicus_corp, "case") <- amicus$case
docvars(amicus_corp, "brief") <- amicus$brief
docvars(amicus_corp, "id") <- amicus$id
docvars(amicus_corp, "brief_party") <- amicus$brief_party

summary(amicus_corp) %>% head()
```

## Create frame dictionary

This dictionary will be used to filter out frame words from paragraphs when counting word frequency.

```{r}
# Remove duplicates
frame_terms <- frame_terms %>%
  select(frame_code, word_clean) %>% 
  unique()

# Get words associated with frames
get_frames <- function(frame){ 
  frame_terms %>%
    filter(frame_code == frame) %>% 
    select(word_clean)
  }

# Initialize list
search_frames <- list()

# Get unique frames
frames <- unique(frame_terms$frame_code)

# Pull words for each frame and save in list
for(i in 1:length(frames)){ 
  search_frames[i] <- get_frames(frames[i])
}

# Name items in list
names(search_frames) <- frames

frame_dict <- dictionary(search_frames)

# Check output
#frame_dict
```

# Look up frames

**Note on arguments in `tokens()` function**: removing punctuation and numbers from terms increases counts. See `amicus-exploring-counting-frames` script to go over evidence.

```{r}
# Tokenize briefs
toks <- tokens(amicus_corp, remove_punct = TRUE, remove_numbers = TRUE)
```

## Create brief summary statistics

The `amicus_corp_summary` data frame contains the summary statistics of all the briefs in our corpus, `amicus_corp`. The one thing it's missing is the number of tokens (i.e. words) in each brief.

```{r}
# Generate summary of amicusCorp
amicus_corp_summary <- summary(object = amicus_corp, 
                      n = nrow(amicus), # default is 100, we need all observations 
                      showmeta = TRUE) %>% # add meta data to summary
  as.data.frame()

# check
tail(amicus_corp_summary)
```

Calculate total number of words in each brief. Join with the amicus_corp_summary

```{r}
# Total number of tokens (before pulling frame words)
nToks <- data.frame(Text = names(ntoken(toks)), 
                     word_ct = ntoken(toks))

# Look at first 6 observations
head(nToks)
```

```{r}
amicus_info <- amicus_corp_summary %>%
  left_join(nToks, by = c("Text")) %>% 
  rename(doc_id=Text)

head(amicus_info)
```

## Generate frame frequencies

Now that we've created our `amicus_info` data frame, it's time to generate the frame frequencies.

(# of words/phrases for a particular frame) / (# of all words in the brief)

```{r}
# Search for frames
amicus_dfm <- tokens_lookup(toks, dictionary = frame_dict, nested_scope = "dictionary",
                         case_insensitive = FALSE, valuetype = "fixed") %>% dfm()

# Generate raw frequencies
amicus_frames_raw <- amicus_dfm %>%
  convert(to = "data.frame") %>%
  left_join(amicus_info, by = "doc_id") %>% 
  janitor::clean_names()

# Check out first 6 observations 
head(amicus_frames_raw, 10)
```

Get frame frequencies - These are calculated by dividing the frame count by the word count.

```{r}
# Generate frame frequencies 
amicus_frames_freq <- amicus_frames_raw %>% 
  pivot_longer(c(abortion_different:womens_rights),
               names_to = "frame", values_to = "frame_ct") %>% 
  mutate(frame_freq = frame_ct/word_ct) %>% 
  select(-frame_ct, -word_ct) %>% 
  pivot_wider(names_from = "frame", values_from = frame_freq)

amicus_frames_freq %>% head(10)
```

## Raw counts + freq

Dataframe with raw counts and frequencies

```{r}
amicus_frames <- amicus_frames_raw %>% 
  select(doc_id:womens_rights, word_ct) %>% 
  left_join(., amicus_frames_freq, 
            by = c("doc_id"), 
            suffix = c("_raw", "_freq")) %>% 
  select(-c(types:sentences)) %>% 
  select(case:brief_party, word_ct, sort(colnames(.))) %>% 
  janitor::clean_names()

amicus_frames %>% head(10)
```

Save

```{r}
#box_write(amicus_frames,
#           file_name = "amicus-level-frame-measures.csv",
#           dir_id = 161199725595)
```

# Case-Level Measures

To generate case-level measures:

1. For each amicus type, sum up the frame frequencies measure for all amicus
2. Divide by the number of briefs for that type
3. Do for all cases

```{r}
amicus_frames <- amicus_frames %>% select(-c(case, brief, brief_party))
```

## Sum up frequencies by amicus type

First, add in amicus type vars

```{r}
sum_amicus <- amicus %>% select(case, id, forgau:sgauo) %>% 
  pivot_longer(., cols = c(forgau:sgauo), names_to = "author_type", values_to = "value") %>% 
  filter(value == 1) %>% 
  left_join(., amicus_frames, by = "id") %>%
  select(case, id, author_type, value, contains("freq"))

sum_amicus %>% head()
```

Then, sum up values by case, amicus type

```{r}
sum_amicus <- sum_amicus %>% 
  select(-value) %>% 
  group_by(case, author_type) %>% 
  summarize(across(contains("freq"), ~ sum(.x, na.rm = TRUE)))

sum_amicus %>% head(5)
```

## Number of briefs per amicus author type per case

Pivot the df and select only rows where value is 1 (indicating that the case has an amicus brief with that author type)

```{r}
case_cts <- amicus %>% select(case, forgau:sgauo) %>% 
  pivot_longer(., cols = c(forgau:sgauo), names_to = "author_type", values_to = "value") %>% 
  filter(value == 1)
```

Then, group by case, type and count the number of rows for each

```{r}
case_cts <- case_cts %>% group_by(case, author_type) %>% summarize(n=n())
case_cts %>% head()
```

## Divide summed frequencies by brief count for final measure

```{r}
case_frames <- sum_amicus %>% left_join(., case_cts, by = c("case", "author_type")) %>% 
  mutate(across(contains("freq"), ~.x/n))
```

Save

```{r}
#box_write(case_frames,
#           file_name = "case-level-frame-measures.csv",
#           dir_id = 161199725595)
```

# Additional Info

```{r, eval=F}
frame = "women's_health"

phrases <- frame_terms %>% 
  filter(frame_code == frame) %>%
  select(word_clean) %>% 
  .[[1]]

# search tokens for the desired phrases, then create DFM
frame_freqs <- toks %>% 
  tokens_compound(pattern = phrase(phrases), concatenator = " ", 
                case_insensitive = FALSE, valuetype = "fixed") %>%
  tokens_select(phrases, case_insensitive = FALSE, valuetype = "fixed") %>%
  dfm()

# convert to df w/one row per brief+counts of phrases, joined w/brief info
frame_word_freqs <- frame_freqs %>% 
  convert(to = "data.frame") %>%
  left_join(amicus_info, by = "doc_id") %>% 
  select(case, id, brief, brief_party, word_ct, everything())

# add amicus author
amicus_type <- amicus %>% select(id, forgauf, legauf, legauo, lifauo,
                                 medauf, medauo, relauf, relauo)
frame_word_freqs <- frame_word_freqs %>% 
  left_join(., amicus_type, on = "id") %>% 
  select(-c(doc_id, legauf, legauo, lifauo, medauo, relauf, relauo)) %>% 
  select(case, brief, id, brief_party, forgauf, medauf, word_ct, sort(colnames(.)))

final <- frame_word_freqs %>% filter(forgauf == 1 | medauf == 1) %>% 
  mutate(case = str_replace(case, 'v\\.', 'v'),
         case = str_replace_all(case, "(,|’|'|\\)|\\()", ""))

decisions <- readxl::read_xlsx("../eda/ab_cases.xlsx", sheet = "dat")
decisions <- decisions %>% 
  mutate(case = str_replace(case, 'v\\.', 'v'),
         case = str_replace_all(case, "(,|’|'|\\)|\\()", ""),
         case = case_when(
           case == "Anders Solicitor of Richland County v Floyd" ~ "Anders v Floyd",
           case == "Ayotte AG of NH v Planned Parenhood of Northern New England" ~ "Ayotte v PP",
           case == "Beal Sec Dept of Public Wel of PA v Doe" ~ "Beal v Doe",
           case == "Bellotti Attorn Gen of MA v Baird 1979" ~ "Bellotti v Baird 1979",
           case == "City of Akron v Akron Center for Rep Health" ~ "City of Akron v Akron Center",
           case == "Colautti Sec of Wel of PA v Franklin" ~ "Colautti v Franklin",
           case == "Doe et al v Bolton Attorn Gen of GA et al" ~ "Doe v Bolton",
           case == "Gonzales AG v Leroy Carhart" ~ "Gonzales v Carhart",
           case == "Gonzales v Planned Parenthood Federation of America" ~ "Gonzales v PP",
           case == "Harris Sec of HHS v McRae" ~ "Harris v McRae",
           case == "Hartigan Attor Gen of IL v David Zbaraz and Allan G Charles" ~ "Hartigan v Zbaraz",
           case == "Hill et al v Colorado" ~ "Hill v Colorado",
           case == "H L v Matheson Gov of UT" ~ "H.L. v Matheson",
           case == "Judy Madsen v Womens Health Center" ~ "Madsen v Womens Health Center",
           case == "Maher Comm of Soc Serv of CT v Roe" ~ "Maher v Roe",
           case == "NOW v Joseph Scheidler 1994" ~ "NOW v Scheidler 1994",
           case == "Ohio v Akron Center for Reproductive Health" ~ "Ohio v Akron Center",
           case == "Planned Parenthood Assoc of Kansas City MO v Ashcroft Attorn Gen of MO" ~
             "PP v Ashcroft",
           case == "Planned Parenhood of SE Penn v Casey" ~ "PP v Casey",
           case == "Planned Parenthood of Central MO v Danforth Attorn Gen of MO" ~ "PP v Danforth",
           case == "Roe et al v Wade District Attorn of Dallas County" ~ "Roe v Wade",
           case == "Rust v Sullivan Sec of HHS" ~ "Rust v Sullivan",
           case == "Joseph Scheidler v NOW 2003" ~ "Scheidler v NOW 2003",
           case == "Scheidler v NOW 2006" ~ "Scheidler v Now 2006",
           case == "Schenck et al v Pro-Choice Network of Western NY" ~
             "Schenck v Pro-Choice Network",
           case == "Stenberg AG of Neb v Leroy Carhart" ~ "Stenberg v Carhart",
           case == "Thornburgh Gov of PA v American College of Obstetricians and Gynecologists" ~
             "Thornburgh v American College of OBGYNs",
           case == "Webster Attor Gen of MO v Reproductive Health Services" ~ "Webster v Reproductive Health Services", 
           case == "Whole Womans Health v Hellerstedt" ~ "WWH v Hellerstedt",
           T ~ case
         )) %>%
  arrange(case)

final %>% 
  janitor::clean_names() %>% 
  left_join(., decisions, by = "case") %>% 
  mutate(death = mortality+morbidity) %>% 
  group_by(fem_win, forgauf, medauf) %>% 
  summarize(n = sum(death))

final %>% 
  janitor::clean_names() %>% 
  left_join(., decisions, by = "case") %>%
  mutate(death = mortality+morbidity,
         health_risk = health_risk+health_risks) %>% 
  filter(forgauf == 1, fem_win == 0) %>% 
  group_by(case, ddate2) %>% 
  summarize(maternal_health = sum(maternal_health), death = sum(death),
            health_risk = sum(health_risk)) %>% 
  arrange(ddate2)

final %>% 
  janitor::clean_names() %>% 
  left_join(., decisions, by = "case") %>%
  mutate(death = mortality+morbidity) %>% 
  filter(medauf == 1, fem_win == 1) %>% 
  arrange(-death) %>% 
  select(case, brief, forgauf, medauf, death, maternal_health)

final %>% 
  janitor::clean_names() %>% 
  left_join(., decisions, by = "case") %>% 
  mutate(illegal_abortion = illegal_abortion+illegal_abortions,
         safe_abortion = safe_abortion + safe_legal_abortion,
         self_abortion = self_abortion+self_abortions+
           self_induced_abortion+self_induced_abortions,
         health_risk = health_risk+health_risks) %>% 
  select(-id, -brief_party, -word_ct, -sentences, -tokens, -types) %>% 
  group_by(fem_win, forgauf, medauf) %>%
  #summarize(across(where(is.numeric), ~sum(.x)))
  summarize(illegal_abortion = sum(illegal_abortion),
            safe_abortion = sum(safe_abortion),
            self_abortion = sum(self_abortion),
            health_risk = sum(health_risk))
```